---
title: "20210204-phia_pca_jsg2145"
author: "Jared Garfinkel"
date: "2/4/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(psych)
library(corrplot)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	# cache = TRUE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# read in data

```{r}
bio <- read_csv(file = "./data/Shims22016adultbio.csv",
                      col_names = TRUE, 
                      col_types = NULL,
                      locale = default_locale(), 
                      na = c("", "NA"), 
                      quoted_na = TRUE,
                      quote = "\"", 
                      comment = "", 
                      trim_ws = TRUE, 
                      skip = 0,
                      n_max = Inf, 
                      progress = show_progress(), 
                      skip_empty_rows = TRUE) %>% 
  janitor::clean_names() %>% 
  mutate(across(.cols = c(-country, -householdid, -personid, -surveystdt, -resultvlc), ~as.numeric(.)))


ind <- read_csv(file = "./data/Shims22016adultind.csv") %>% 
  janitor::clean_names() %>% 
  mutate(across(.cols = c(-country, -householdid, -personid, -surveystdt), ~as.numeric(.x)))
```

# bt_status = 1, hivstatusfinal = 1, arvstatus = 1

## This is the primary difference between this data set and others

```{r}
bio_select <- bio %>%
    left_join(ind, by = c("personid", "country", "householdid", "bt_status", "gender", "age", "surveystdt", "hivstatusfinal", "varstrat", "varunit")) %>%
  # "avdprgmeth", "hivtstnors", "hivpostell", "sptgrgt", "stdl", "seekhelp", "unwantsxhelp")) %>%
    select(-contains("vmpstw"), -contains("hivk"), -contains("casi"), -contains("intw")) %>%
    filter(bt_status == 1,
           hivstatusfinal == 1) %>%
           # arvstatus == 1) %>%
    select(personid, age, gender, contains(c("status", "aware", "art", "vl", "tri90", "arv", "test", "hiv"))) %>%
    mutate( resultvlc = recode(resultvlc,
                    "< LLOD" = "1",
                    "< LLOQ: 20" = "20",
                    "< LLOQ: 40" = "40",
                    "< LLOQ: 400" = "400",
                    "< LLOQ: 839" = "839",
                    "> ULOQ 10000000" = "10000000"),
            resultvlc = as.numeric(resultvlc),
            vlunder200 = if_else(resultvlc < 200, 1, 2))

bio_select %>% 
  select(vlunder200, everything()) %>% 
  group_by(vlunder200) %>%
  summarize(freq = n())
```

# clean the data

```{r}
recode_one_zero <- function(x) {
  if_else( is.na(x), 99,
  if_else( as.numeric(x) %in% c(0, 2), 0, # JSG
  if_else( as.numeric(x) == 1, 1, 99)
  ))
}

bio_select_mod <- bio_select %>%
    mutate( reported_positive = if_else( hivrslr == 1 | hivtstrslt == 1 | hivpsbp == 1 | hivrtpg == 1 | hivtfposy > 0, 1, 0, missing = 0),
            never_tested = if_else(hivtstever == 2, 1, 0, missing = 0),
            reported_negative = if_else(hivpsbp == 2 | (hivtsbp = 1 & hivrtpg == 2) | hivrslr == 2 | hivtstrslt == 2, 1, 0, missing = 0),
            pos_test_year = if_else(hivtfposy > 0, 1, 0),
            tested12mo = if_else(testedreceiveddetail %in% c(1,4), 1, 0),
            cd4testever = recode_one_zero(cd4testever),
            arvstakenev = if_else(arvstakenev == 1, 1, 0, missing = 0),
            arvscurrent = if_else(arvscurrent == 1, 1, 0, missing = 0),
            oth_art_care = if_else(arvtklb == 1 | arvcntn == 1 | hivcare == 1, 1, 0, missing = 0),
            across(-personid, recode_one_zero))

```

# Keep "arvstatus" in the model

```{r}
modelvars <- c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "arvstatus", "vlunder200")
# Remove all rows which have values other than 0 or 1
aware_patterns <- bio_select_mod %>% 
  dplyr::select( all_of(modelvars) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1))) %>% 
  mutate(across(.cols = c(reported_negative, never_tested), .fns = ~recode(., "0" = 1, "1" = 0)))
```

# eigenvalue scree plot

```{r}
awareparallel = fa.parallel(aware_patterns, cor = "tet", fa = "both")
awareparallel$pc.values
awareparallel$pc.sim
```

# cronbach's alpha

```{r}
awarealpha = alpha(aware_patterns)
awarealpha
```

This indicates that "never tested" might should be dropped?

If never_tested is dropped, 

# Exploratory Factor Analysis

```{r}
# ?fa

res = fa(aware_patterns, 2, cor = 'tet', rotate = "geominQ", fm = "wls")
# res3 = fa(aware_patterns, 3, cor = 'tet', rotate = "geominQ", fm = "wls")

res
# res3
```

```{r}
fa.diagram(res, digits = 2)
```

# correlation analysis

```{r}
M = cor(aware_patterns)
corrplot(M, type = c("upper"), order = c("hclust"))
```

## Extract correlation values

```{r}
output = vector(mode = "list", length = ncol(M))
rname = NULL
cname = NULL
for(i in 1:nrow(M)) {
  for(j in 1:ncol(M)) {
    rname[[i]] = dimnames(M)[[c(1, i)]]
    cname[[j]] = dimnames(M)[[c(2, j)]]
    output[[c(i, j)]] = tibble_row(i, j, rname[[i]], cname[[j]], M[[i, j]])
  }
}
```

```{r, result = "hide"}

matrify = function(pdata = output) {
  for(i in 1:46) {
      result[[i]] = tibble(pdata[[c(i, i)]])
  }
  result = bind_rows(result)
  return(result)
}
```

```{r}
matrify2 = function(pdata = output) {
  result = NULL
  for(i in 1:length(pdata)) {
      result[[i]] = tibble(pdata[[i]])
  }
  result = bind_rows(result)
  return(result)
}

df2 = vector(mode = "list", length = ncol(M))
for(j in 1:ncol(M)) {
  df2[[j]] = matrify2(pdata = output[[j]])
}
```

```{r}
union_df = bind_rows(df2) %>% 
  rename("value" = "M[[i, j]]",
         "column" = "cname[[j]]",
         "row" = "rname[[i]]")
```

```{r high df}
high_df = union_df %>% 
  filter(abs(value) >= 0.7,
         value != 1,
         i < j)

high_df
```

# principal in psych package

```{r principal}
fit = principal(M, 4, cor = "tet", rotate = "geominQ")
fit
```

# iclust in psych package

```{r iclust}
# ?iclust
# pdf()
fit2 = iclust(M, plot = TRUE)
# dev.off()
fit2
```

## redo the pca

```{r}
pca_df = bio %>% 
    left_join(ind, by = c("personid", "country", "householdid", "bt_status", "gender", "age", "surveystdt", "hivstatusfinal", "varstrat", "varunit")) %>% 
  # "avdprgmeth", "hivtstnors", "hivpostell", "sptgrgt", "stdl", "seekhelp", "unwantsxhelp")) %>%
    select(-contains("vmpstw"), -contains("hivk"), -contains("casi"), -contains("intw")) %>%
    filter(bt_status == 1,
           hivstatusfinal == 1) %>%
           # arvstatus == 1) %>% 
    select(personid, gender, contains(c("status", "aware", "art", "vl", "tri90", "arv", "test", "hiv")), -contains(c("part", "start", "partner", "vlnc", "child", "postell", "nors", "lifetime", "hivccost", "arvnrpg")), -ends_with(c("ty", "sy", "fy", "cly"))) %>% #remove age
    mutate( resultvlc = recode(resultvlc, 
                    "< LLOD" = "1",
                    "< LLOQ: 20" = "20",
                    "< LLOQ: 40" = "40",
                    "< LLOQ: 400" = "400",
                    "< LLOQ: 839" = "839",
                    "> ULOQ 10000000" = "10000000"),
            resultvlc = as.numeric(resultvlc),
            vlunder200 = if_else(resultvlc < 200, 1, 2)) %>% 
  select(-resultvlc, -indstatus, -bt_status, -hivstatusfinal, -hiv1statusfinalsurvey)
  # mutate(across(.fns = ~na_if(., 99)))
  # mutate(across(.fns = ~replace_na(., NA)))
  data.frame()
  # mutate(across(.fns=as_factor)) # took out hivtfposy for categorical data reasons, hivrslr, arvtklb, arvcntn missing > 0.95
```

```{r}
output = NULL
for (i in 1:72) {
  output[[i]] = unique(pull(pca_df[,-1], i))
  output[[i]] = rbind(output[[i]])
}

output %>% 
  unlist() %>% 
  unique()
output_df = bind_rows(output)
output

unique(pca_df[[6]])


```

```{r}

principal(pca_df[,-1], nfactors = 6)
which(skimr::skim(pca_df)$complete_rate == 1)

pca_df_skimd = pca_df[,-c(9, 25, 26, 32:47, 50, 54, 58:66, 72:86)]

pca_df_skimd = pca_df_skimd %>% 
  select(-personid)

pca_df_skimd_df = pca_df_skimd %>% 
  mutate(across(.fns = as.numeric),
         across(.fns = ~replace_na(0)))

# pca_df_skimd[37]

skimr::skim(pca_df[,c(2, 8, 22, 25, 70)])

pca_df[3]

pca_df %>% 
  str()

pca_df %>% 
  pull(6)
?fa.parallel
?fa
?cor
corrplot(pca_df[,-1])
M = cor(pca_df[,-1])
corrplot(M)

scree = fa(pca_df[,-1], nfactors = 7)
which(complete.cases(pca_df))
# scree = principal(pca_df[,-1], nfactors = 7)
# scree = fa.parallel(pca_df[,c(2, 8, 22, 25, 70)], cor = "mixed", fm = "wls", sim = FALSE)
# scree = fa.parallel(pca_df[,-1], fm = "wls")
```

```{r}
pca_df2 <- bio %>%
    left_join(ind, by = c("personid", "country", "householdid", "bt_status", "gender", "age", "surveystdt", "hivstatusfinal", "varstrat", "varunit")) %>%
  # "avdprgmeth", "hivtstnors", "hivpostell", "sptgrgt", "stdl", "seekhelp", "unwantsxhelp")) %>%
    select(-contains("vmpstw"), -contains("hivk"), -contains("casi"), -contains("intw")) %>%
    filter(bt_status == 1,
           hivstatusfinal == 1) %>%
           # arvstatus == 1) %>%
    select(personid, age, gender, contains(c("status", "aware", "art", "vl", "tri90", "arv", "test", "hiv"))) %>%
    mutate( resultvlc = recode(resultvlc,
                    "< LLOD" = "1",
                    "< LLOQ: 20" = "20",
                    "< LLOQ: 40" = "40",
                    "< LLOQ: 400" = "400",
                    "< LLOQ: 839" = "839",
                    "> ULOQ 10000000" = "10000000"),
            resultvlc = as.numeric(resultvlc),
            vlunder200 = if_else(resultvlc < 200, 1, 2)) %>% 
  select(hivrslr, hivtstrslt, hivpsbp, hivrtpg, hivtstever, hivtsbp, cd4testever, arvstakenev, arvscurrent, arvtklb, arvcntn, hivcare)
```

```{r}

skimr::skim(pca_df2)
?complete.cases

fit = fa(pca_df2, nfactors = 2, fm = "wls")

# output = NULL
# for (i in 1:nrow(pca_df2)) {
#   output[[i]] = complete.cases(pca_df2[i,])
# }

pca_df2 %>% 
  mutate(complete = complete.cases(.)) %>% 
  select(complete, everything()) %>% 
  filter(complete == TRUE)

```

```{r}
skimr::skim(bio_select)
```

# remove never_tested, reported_negative

```{r}
modelvars <- c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "arvstatus", "vlunder200")
# Remove all rows which have values other than 0 or 1
aware_patterns <- bio_select_mod %>% 
  dplyr::select( all_of(modelvars) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1))) %>% 
  mutate(across(.cols = c(reported_negative, never_tested), .fns = ~recode(., "0" = 1, "1" = 0)))
```

# eigenvalue scree plot

```{r}
awareparallel = fa.parallel(aware_patterns, cor = "tet", fa = "both")
awareparallel$pc.values
awareparallel$pc.sim
```

```{r}
?fa

res2 = fa(aware_patterns, 4, cor = 'tet', rotate = "geominQ", fm = "wls")
# res3 = fa(aware_patterns, 3, cor = 'tet', rotate = "geominQ", fm = "wls")

res2
# res3

fa.diagram(res2, digits = 2)
```