---
title: "phia_lca5_jsg2145"
author: "Jared Garfinkel"
date: "8/28/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(randomLCA)
library(gt)
library(caret)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	cache = TRUE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## read in data

```{r, echo = FALSE, results = 'hide', cache = FALSE, warning = FALSE}
# Load in PHIA data

recode_one_zero <- function(x) {
  if_else( is.na(x), 99,
  if_else( as.numeric(x) %in% c(0, 2), 0, # JSG
  if_else( as.numeric(x) == 1, 1, 99)
  ))
}

## read in data
bio <- read_csv("C:\\Users\\jared\\OneDrive\\Mailman\\Summer 2020\\PHIA ICAP\\zimphia2015adultbio.csv") %>% 
  janitor::clean_names() %>% 
  mutate(across(.cols = c(-country, -householdid, -personid, -centroidid, -surveystdt, -resultvlc), ~as.numeric(.x)))
ind <- read_csv("C:\\Users\\jared\\OneDrive\\Mailman\\Summer 2020\\PHIA ICAP\\zimphia2015adultind.csv") %>% 
  janitor::clean_names() %>% 
  mutate(across(.cols = c(-country, -householdid, -personid, -centroidid, -surveystdt), ~as.numeric(.x)))

bio_select <- bio %>% 
    left_join(ind, by = c("personid", "country", "householdid", "centroidid", "bt_status", "gender", "age", "surveystdt", "hivstatusfinal", "varstrat", "varunit")) %>% 
  # "avdprgmeth", "hivtstnors", "hivpostell", "sptgrgt", "stdl", "seekhelp", "unwantsxhelp")) %>%
    select(-contains("vmpstw"), -contains("hivk"), -contains("casi"), -contains("intw")) %>%
    filter(bt_status == 1,
           hivstatusfinal == 1) %>% 
    select(personid, age, gender, contains(c("status", "aware", "art", "vl", "tri90", "arv", "test", "hiv"))) %>% 
    mutate( resultvlc = recode(resultvlc, 
                               "ND" = "1",
                               "NQ: < 20" = "20",
                               "NQ: < 182" = "182"),
            resultvlc = as.numeric(resultvlc),
            vlunder200 = if_else(resultvlc < 200, 1, 2, missing = 2)) # JSG (from Stephen)

#table(vls = bio_select$vls, vl_lt_200 = bio_select$vl_lt_200)
# head(ind)
# head(bio)
# head(bio_select)
# str(ind$bt_status)
# bio$varstrat
# bio_select %>%
#   select(ends_with("y"))
# bio_select %>%
#   group_by(resultvlc) %>%
#   nest() %>%
#   filter(str_detect(resultvlc, "[:alpha:]"))
# which(is.na(bio_select$resultvlc == TRUE))
# bio_select$resultvlc
```


```{r}
bio_select_mod <- bio_select %>%
    mutate( reported_positive = if_else( hivrslr == 1 | hivtstrslt == 1 | hivpsbp == 1 | hivrtpg == 1 | hivtfposy > 0, 1, 0, missing = 0),
            never_tested = if_else(hivtstever == 2, 1, 0, missing = 0),
            reported_negative = if_else(hivpsbp == 2 | (hivtsbp = 1 & hivrtpg == 2) | hivrslr == 2 | hivtstrslt == 2, 1, 0, missing = 0),
            pos_test_year = if_else(hivtfposy > 0, 1, 0),
            tested12mo = if_else(testedreceiveddetail %in% c(1,4), 1, 0),
            cd4testever = recode_one_zero(cd4testever),
            arvstakenev = if_else(arvstakenev == 1, 1, 0, missing = 0),
            arvscurrent = if_else(arvscurrent == 1, 1, 0, missing = 0),
            oth_art_care = if_else(arvtklb == 1 | arvcntn == 1 | hivcare == 1, 1, 0, missing = 0),
            across(-personid, recode_one_zero))

bio_select_var = bio %>% 
  filter(bt_status == 1,
         hivstatusfinal == 1) %>% 
  select(personid, awareartselfreported)

# nrow(bio_select_var) # 3507
```

## LCA, abridged

```{r}
modelvars <- c( "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "vlunder200")
# Remove all rows which have values other than 0 or 1
aware_patterns <- bio_select_mod %>% 
  select( all_of(modelvars) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
awareARTVL.lca3 <- randomLCA(aware_patterns, nclass = 3, quadpoints = 30)

outcomep <- round(awareARTVL.lca3$outcomep, 4)
colnames(outcomep) <- names(aware_patterns)

OutcomeProbs.Table.AwareART <- as.data.frame( cbind( Class = c("Class 1 - Unaware", "Class 2 - Aware, not on ART", "Class 3 - Aware, on ART"), outcomep)) %>%
  mutate(across(-Class, as.numeric))
OutcomeProbs.Table.AwareART
```

```{r}
transpose_class = t(outcomep) %>% 
  data.frame() %>% 
  tibble(Predictor = row.names(.)) %>% 
  select(Predictor, everything()) %>% 
  rename("Class 1 Aware - On Art" = "X1",
         "Class 2 Aware - Not on ART" = "X2",
         "Class 3 Unaware" = "X3")
gt(transpose_class)
```


## extract classes

```{r}
set.seed(719)
lca_out_full = postClassProbs(awareARTVL.lca3) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3, 1,
                                 if_else(class_3 > class_1, 3, 1)), 
                         if_else(class_2 > class_3, 2, 3)),
         vlsaware = if_else(class == 1, 1, 0), # coded "forwards" meaning 1 is aware on art, 0 is EITHER unaware OR "aware and not on art"
         awarenotart = if_else(class == 2, 0, 1), # coded "backwards" meaning 0 is "aware not on art", 1 is not "aware not on art", or in other words, neither unaware nor aware and not on art - in terms of  class, this means that an individual may be EITHER "aware AND on art" OR unaware
         unaware = if_else(class == 3, 0, 1)) # coded "backwards" meaning 0 is unaware, 1 is not unawarecoded "forwards" meaning 1 is aware on art, 0 is EITHER unaware OR "aware and not on art"
```

```{r}
full_dat_out = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_full, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "vlunder200"))

```

## awareartselfreported

The agreement between variables is uncanny

```{r}
xtabs(~awareartselfreported.y + class, data = full_dat_out, exclude = "99")
```

## awareness

```{r}
xtabs(~aware + unaware, data = full_dat_out, exclude = "99")
```


```{r}
sensitbl = tibble(name = c("Sensitivity", "Specificity", "PPV", "NPV"),
                  value = c(round(2667/2667, 2), round(714/839, 2), round(2667/2792, 2), round(714/714)))
gtbl = gt(sensitbl)


gt_tbl <- 
  gtbl %>% 
  # gt(rowname_col = "Sensitivity")
  tab_row_group(group = "doctor",
                rows = c(1:2)) %>%
  tab_row_group(group = "patient",
                rows = c(3:4))

gt_tbl
```

## on ART

```{r}
xtabs(~ art + vlsaware, data = full_dat_out, exclude = "99") %>% 
  ftable(.)
```

```{r}
vlsupprestbl = tibble(name = c("Sensitivity", "Specificity", "PPV", "NPV"),
                  value = c(round(2349/2349, 2), round(1008/1157, 2), round(2349/2498, 2), round(1008/1008, 2)))
gtbl = gt(vlsupprestbl)


gt_tbl <- 
  gtbl %>% 
  # gt(rowname_col = "Sensitivity")
  tab_row_group(group = "doctor",
                rows = c(1:2)) %>%
  tab_row_group(group = "patient",
                rows = c(3:4))

gt_tbl
```

## Aware - not on ART

If a person is unaware, then they would not be on art. These should not be included in the "aware - not on art" variable. So, we take those out using this table. A "0" in unaware means they were unaware. A "1" in unaware means they were not unaware. Only those who were not unaware should be counted.

```{r}
xtabs(~ art + unaware + awarenotart, data = full_dat_out, exclude = "99") %>% 
  ftable(col.vars = c(2:3))
```

```{r}
notonartbl = tibble(name = c("Sensitivity", "Specificity", "PPV", "NPV"),
                  value = c(round(2349/2349, 2), round(294/318, 2), round(2349/2373, 2), round(294/294, 2)))
gtbl = gt(notonartbl)


gt_tbl <- 
  gtbl %>% 
  # gt(rowname_col = "Sensitivity")
  tab_row_group(group = "doctor",
                rows = c(1:2)) %>%
  tab_row_group(group = "patient",
                rows = c(3:4))

gt_tbl
```

# Calculating Prevalence using confusionMatrix()

```{r Diggle, include = FALSE, eval = FALSE}
prevalence.bayes <- function(theta, 
                             T, 
                             n, 
                             lowse=0.5, 
                             highse=1.0, 
                             sea=1, 
                             seb=1, 
                             lowsp=0.5, 
                             highsp=1.0, 
                             spa=1, 
                             spb=1, 
                             ngrid=20, 
                             coverage=0.95) {
  ibeta <- function(x,a,b) {
    pbeta(x, a, b) * beta(a, b)
  }
  ntheta <- length(theta)
  bin.width <- (theta[ntheta] - theta[1])/(ntheta - 1)
  theta <- theta[1] + bin.width*(0:(ntheta - 1))
  integrand <- array(0, c(ntheta, ngrid, ngrid))
  h1 <- (highse - lowse)/ngrid
  h2 <- (highsp - lowsp)/ngrid
  for (i in 1:ngrid) {
    se <- lowse + h1*(i - 0.5)
    pse <- (1/(highse - lowse))*dbeta((se - lowse)/(highse - lowse), sea, seb)
    for (j in 1:ngrid) {
      sp <- lowsp + h2*(j - 0.5)
      psp <- (1/(highsp - lowsp))*dbeta((sp - lowsp)/(highsp - lowsp), spa, spb)
      c1 <- 1 - sp
      c2 <- se + sp - 1
      f <- (1/c2)*choose(n, T)*(ibeta(c1 + c2, T + 1, n - T + 1) - ibeta(c1, T + 1, n - T + 1))
      p <- c1 + c2*theta
      density <- rep(0, ntheta)
      for (k in 1:ntheta) {
        density[k] <- dbinom(T, n, p[k])/f
      }
      integrand[,i,j] <- density*pse*psp
    }
  }
  post <- rep(0, ntheta)
  for (i in 1:ntheta) {
    post[i] <- h1*h2*sum(integrand[i,,])
  }
  ord <- order(post, decreasing = T)
  mode <- theta[ord[1]]
  take <- NULL
  prob <- 0
  i <- 0
  while ((prob < coverage/bin.width) & (i < ntheta)) {
    i <- i + 1
    take <- c(take, ord[i])
    prob <- prob + post[ord[i]]
  }
  if (i == ntheta) {
    print("WARNING: range of values of theta too narrow")
  }
  interval <- theta[range(take)]
  list(theta = theta, post = post, mode = mode, interval = interval, coverage = prob*bin.width)
}
```

```{r, include = FALSE, eval = FALSE}
stable = function(probs = probs2){
    
  class <- if_else(probs[[1]]$Outcome[1] > probs[[1]]$Outcome[2], 
                         if_else(probs[[1]]$Outcome[2] > probs[[1]]$Outcome[3], 1,
                                 if_else(probs[[1]]$Outcome[3] > probs[[1]]$Outcome[1], 3, 1)), 
                         if_else(probs[[1]]$Outcome[2] > probs[[1]]$Outcome[3], 2, 3))
  
  aware <- if_else(class %in% c(1, 2), 1, 0)
  
  art <- if_else(class == 1, 1, 0)
  
  unaware <- if_else(class == 3, 1, 0)
  
  sens <- apply(probs[[class]], 1, function(x)
    sprintf("%3.2f (%3.2f, %3.2f)", 
            x[1], 
            x[2], 
            x[3])
  )
  
  spec <- apply(probs[[aware]], 1, function(x)
    sprintf("%3.2f (%3.2f, %3.2f)", 
            1 - x[1], 
            1 - x[3], 
            1 - x[2])
  )
  
  stable <- data.frame(sens, spec)
  
  names(stable) <- c("Sensitivity", "Specificity")
  
  return(list(aware, art, unaware))
}
```

## calculate confidence intervals for sensitivity and specifity to calculate prevalence

### Awareness

```{r}
conmat_aware_df = full_dat_out

conmat_aware_df$class = factor(conmat_aware_df$class, levels = c("aware - on art" = 1, "aware - not on art" = 2, "unaware" = 3), labels = c("aware", "aware", "unaware"))
conmat_aware_df$aware = factor(conmat_aware_df$aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(conmat_aware_df$class, conmat_aware_df$aware, positive = "aware")
cMat

```

```{r from Diggle, include = FALSE, eval = FALSE}
c = max(0, (a - (1 - sp))/(se + sp - 1))
d = min(1, (b - (1 - sp))/(se + sp - 1))
theta = paste(c, ", ", d)
theta
```

```{r, include = FALSE, eval = FALSE}
a = cMat$overall["AccuracyLower"]
b = cMat$overall["AccuracyUpper"]
se = cMat$byClass["Sensitivity"]
sp = cMat$byClass["Specificity"]
```

### On ART

```{r}
conmat_art_df = full_dat_out

conmat_art_df$class = factor(conmat_art_df$class, levels = c("aware - on art" = 1, "aware - not on art" = 2, "unaware" = 3), labels = c("on art", "not on art", "not on art"))
conmat_art_df$art = factor(conmat_art_df$art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat.art = caret::confusionMatrix(conmat_art_df$class, conmat_art_df$art, positive = "on art")
cMat.art
```


