---
title: "phia_lca2_jsg2145"
author: "Jared Garfinkel"
date: "8/9/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(randomLCA)
library(pROC)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r read in data}
# Read public-release datasets
#Biomarker dataset
biomarker <- read_csv(file = "./data/Shims22016adultbio.csv",
                      col_names = TRUE, 
                      col_types = NULL,
                      locale = default_locale(), 
                      na = c("", "NA"), 
                      quoted_na = TRUE,
                      quote = "\"", 
                      comment = "", 
                      trim_ws = TRUE, 
                      skip = 0,
                      n_max = Inf, 
                      progress = show_progress(), 
                      skip_empty_rows = TRUE)
```

```{r clean data}
bio_dat <- biomarker %>% 
  filter(hiv1statusfinalsurvey == 1) %>%
  select(personid, 
         hiv1statusfinalsurvey,
         awareselfreported,
         gender,
         arvstatus, 
         artselfreported, 
         resultvlc, 
         vls, 
         tri90, 
         tri90aware, 
         tri90art, 
         tri90vls) %>%
  mutate(resultvlc = recode(resultvlc, 
                    "< LLOD" = "1",
                    "< LLOQ: 20" = "20",
                    "< LLOQ: 40" = "40",
                    "< LLOQ: 400" = "400",
                    "< LLOQ: 839" = "839",
                    "> ULOQ 10000000" = "10000000"),
         resultvlc = as.numeric(resultvlc),
         vlunder200 = if_else(resultvlc < 200, 1, 2),
         across(.funs = as_factor))
```

```{r}
str(bio_dat)
```


```{r}
# aware:                1 - Aware or considered aware because ARVs detectable 
#                       2 - Unaware and ARVs not detectable, or unaware and ARV testing results missing 
#                       99 - Missing
#
# art:                  1 - ARVs detectable, self-reported on ART, or both ARVs detectable and self-reported on ART 
#                       2 - Unaware or aware, ARVs not detectable and self-reported not on ART, or aware, missing ARV testing data and self-reported not on ART 
#                       99 - Missing
# awareselfreported:    1 - Self-report aware of HIV + status 
#                       2 - Self-report not aware of HIV + status 
#                       99 - missing
# artselfreported:      1 - On ART 
#                       2 - Not on ART 
#                       99 - Missing
# awareartselfreported: 1 - Self-report as not previously diagnosed 
#                       2 - Self-report as previously diagnosed, not on ART 
#                       3 - Previously diagnosed, on ART 
#                       99 - Missing, including incomplete tri90 information
# arvstatus:            1 - ARV detected 
#                       2 - ARV not detected 
#                       99 - Missing
# resultvlc:            > ULOQ 10000000 - Upper limit of quantification 10000000
#                       < LLOD - less than lower limit of detection 
#                       < LLOQ: 839 - less than lower limit of quantification of 839 
#                       < LLOQ: 400 - less than lower limit of quantification of 400
#                       < LLOQ: 40 - less than lower limit of quantification of 40 
#                       < LLOQ: 20 - less than lower limit of quantification of 20
# hivselfreport:        1 - Self-reported positive 
#                       2 - Self-reported negative 
#                       3 - Self-reported never tested or never received test result 
#                       99 - Missing
# tri90art              1 - ARVs detectable, self-reported on ART, or both ARVs detectable and self-reported on ART 
#                       2 - ARVs not detectable and self-reported not on ART or missing ARV data and self-reported not on ART 
#                       3 - Recoded as not on ART (unaware and ARVs not detectable or unaware and ARVs missing) 
#                       99 - Incomplete Tri90 information
#
```

```{r clean data for lca}
bio_dat_lca = bio_dat %>% 
  select(awareselfreported, 
         arvstatus, 
         artselfreported, 
         gender,
         vls, 
         tri90, 
         tri90aware, 
         tri90art, 
         tri90vls,
         vlunder200) %>% 
  mutate(awareselfreported = recode(awareselfreported, "1" = "1", "2" = "0", "99" = NULL),
         arvstatus = recode(arvstatus, "1" = "1", "2" = "0", "99" = NULL),
         artselfreported = recode(artselfreported, "1" = "1", "2" = "0", "99" = NULL),
         gender = recode(gender, "1" = "1", "2" = "0", "99" = NULL),
         vls = recode(vls, "1" = "1", "2" = "0", "99" = NULL),
         tri90 = recode(tri90, "1" = "1", "2" = "0", "99" = NULL),
         tri90aware = recode(tri90aware, "1" = "1", "2" = "0", "99" = NULL),
         tri90art = recode(tri90art, "1" = "1", "2" = "0", "99" = NULL),
         tri90vls = recode(tri90vls, "1" = "1", "2" = "0", "99" = NULL),
         vlunder200 = recode(vlunder200, "1" = "1", "2" = "0", "99" = NULL))
```

```{r simplify code}
bio_dat_lca_full = bio_dat %>% 
  select(awareselfreported, 
         arvstatus, 
         gender,
         vlunder200) %>% 
  mutate(across(everything(), ~recode(.x, "1" = "1", "2" = "0", "99" = NULL)))
```


```{r try a function, eval = FALSE}
bio_dat1 = bio_dat %>% 
  mutate(awareselfreported = recode(awareselfreported, "1" = "1", "2" = "0", "99" = NULL))

lca_prepper = function(.) {
  mutate(across(.funs = recode(., "1" = "1", "2" = "0", "99" = NULL))
}
```


# "Full Model"

```{r}
# full model
bio_lca_full = bio_dat_lca %>% 
  select(awareselfreported, arvstatus, vlunder200)



set.seed(22)
bio_lca = randomLCA::randomLCA(bio_lca_full, calcSE = TRUE)
summary(bio_lca)

probs = outcomeProbs(bio_lca)
postClassProbs(bio_lca)
BIC(bio_lca)
```

```{r code for sens and spec, results = "hide"}
diseased <- ifelse(probs[[1]]$Outcome[1] < probs[[2]]$Outcome[1], 2, 1)
notdiseased <- 3 - diseased
sens <- apply(probs[[diseased]], 1, function(x)
  
  sprintf("%3.2f (%3.2f, %3.2f)", 
          x[1], 
          x[2], 
          x[3])
)
spec <- apply(probs[[notdiseased]], 1, function(x)
  sprintf("%3.2f (%3.2f, %3.2f)", 
          1 - x[1], 
          1 - x[3], 
          1 - x[2])
)

stable <- data.frame(sens, spec)
names(stable) <- c("Sensitivity", "Specificity")
print(stable, row.names = TRUE)

```

```{r stable function}
stable = function(probs = probs){
    
  diseased <- ifelse(probs[[1]]$Outcome[1] < probs[[2]]$Outcome[1], 2, 1)
  
  notdiseased <- 3 - diseased
  
  sens <- apply(probs[[diseased]], 1, function(x)
    sprintf("%3.2f (%3.2f, %3.2f)", 
            x[1], 
            x[2], 
            x[3])
  )
  
  spec <- apply(probs[[notdiseased]], 1, function(x)
    sprintf("%3.2f (%3.2f, %3.2f)", 
            1 - x[1], 
            1 - x[3], 
            1 - x[2])
  )
  
  stable <- data.frame(sens, spec)
  
  names(stable) <- c("Sensitivity", "Specificity")
  
  return(print(stable, row.names = TRUE))
}

# stable(probs2)
```

# "Fuller model"

```{r}
# full model
set.seed(22)
bio_lca_fuller = randomLCA::randomLCA(bio_dat_lca_full, random = FALSE, calcSE = TRUE)

probs_full = outcomeProbs(bio_lca_fuller)

stable(probs_full)
```

# "Simple Model"

A model of sensitivity and specificity for just one variable should be compared to a true (known) outcome. As such, a contingency table should be sufficient to calculate the sensitivity and specificity of each variable.

## awareselfreported

```{r}
xtabs(~ awareselfreported + tri90aware, data = bio_dat_lca)
```

Assuming the final classifications are true for tri90aware, there are 59 false negatives and no false positives. This corresponds to a sensitivity of 2591/2650, or `r round(2591/2650, 3)`, and a specificity of 346/ 346, or 1.00. The prevalance (of awareness of HIV status) is therefore `r round(2650/2996, 3)`. The PPV can be calculated as 2591/2591, or 1.00, and the NPV is 346/405 = `r round(346/405, 3)`.

## arvstatus

```{r}
xtabs(~ arvstatus + tri90aware, data = bio_dat_lca)
```

Assuming the final classifications are true for tri90aware, there are 369 false negatives and no false positives. This corresponds to a sensitivity of 2279/2648, or `r round(2279/2648, 3)`, and a specificity of 345/345, or 1.00. The prevalance (of awareness of HIV status) is therefore `r round(2648/2993, 3)`. The PPV can be calculated as 2279/2279, or 1.00, and the NPV is 345/714 = `r round(345/714, 3)`.

## vlunder200

```{r}
xtabs(~ vlunder200 + tri90aware, data = bio_dat_lca)
```

Assuming the final classifications are true for tri90aware, there are 546 false negatives and 16 false positives. This corresponds to a sensitivity of 2105/2651, or `r round(2105/2651)`, and a specificity of 330/346, or `r round(330/346, 3)`. The prevalance (of awareness of HIV status) is therefore `r round(2651/2997, 3)`. The PPV can be calculated as 2105/2121, or `r round(2105/2121, 3)`, and the NPV is 330/876 = `r round(330/876, 3)`.

# ROC

```{r pROC}
roc(bio_dat$tri90aware, c(bio_dat$awareselfreported + bio_dat$arvstatus + bio_dat$vlunder200), data = bio_dat) %>% 
  plot()
test1 = roc(tri90aware ~ awareselfreported, bio_dat)
test2 = roc(tri90aware ~ arvstatus, bio_dat)
roc.test(test1, test2)

roc.test(bio_dat$tri90aware, bio_dat$awareselfreported, bio_dat$arvstatus)
```


