---
title: "phia_lca8_jsg2145"
author: "Jared Garfinkel"
date: "9/11/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(randomLCA)
library(gt)
library(caret)
knitr::opts_chunk$set(
	echo = FALSE,
	warning = FALSE,
	message = FALSE,
	cache = TRUE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
bio <- read_csv(file = "./data/Shims22016adultbio.csv",
                      col_names = TRUE, 
                      col_types = NULL,
                      locale = default_locale(), 
                      na = c("", "NA"), 
                      quoted_na = TRUE,
                      quote = "\"", 
                      comment = "", 
                      trim_ws = TRUE, 
                      skip = 0,
                      n_max = Inf, 
                      progress = show_progress(), 
                      skip_empty_rows = TRUE) %>% 
  janitor::clean_names() %>% 
  mutate(across(.cols = c(-country, -householdid, -personid, -surveystdt, -resultvlc), ~as.numeric(.x)))


ind <- read_csv(file = "./data/Shims22016adultind.csv") %>% 
  janitor::clean_names() %>% 
  mutate(across(.cols = c(-country, -householdid, -personid, -surveystdt), ~as.numeric(.x)))
```

```{r}
bio_select <- bio %>% 
    left_join(ind, by = c("personid", "country", "householdid", "bt_status", "gender", "age", "surveystdt", "hivstatusfinal", "varstrat", "varunit")) %>% 
  # "avdprgmeth", "hivtstnors", "hivpostell", "sptgrgt", "stdl", "seekhelp", "unwantsxhelp")) %>%
    select(-contains("vmpstw"), -contains("hivk"), -contains("casi"), -contains("intw")) %>%
    filter(bt_status == 1,
           hivstatusfinal == 1) %>% 
    select(personid, age, gender, contains(c("status", "aware", "art", "vl", "tri90", "arv", "test", "hiv"))) %>% 
    mutate( resultvlc = recode(resultvlc, 
                    "< LLOD" = "1",
                    "< LLOQ: 20" = "20",
                    "< LLOQ: 40" = "40",
                    "< LLOQ: 400" = "400",
                    "< LLOQ: 839" = "839",
                    "> ULOQ 10000000" = "10000000"),
            resultvlc = as.numeric(resultvlc),
            vlunder200 = if_else(resultvlc < 200, 1, 2)) 

bio_select %>% 
  select(vlunder200, everything()) %>% 
  group_by(vlunder200) %>%
  summarize(freq = n())
```

```{r}

recode_one_zero <- function(x) {
  if_else( is.na(x), 99,
  if_else( as.numeric(x) %in% c(0, 2), 0, # JSG
  if_else( as.numeric(x) == 1, 1, 99)
  ))
}

bio_select_mod <- bio_select %>%
    mutate( reported_positive = if_else( hivrslr == 1 | hivtstrslt == 1 | hivpsbp == 1 | hivrtpg == 1 | hivtfposy > 0, 1, 0, missing = 0),
            never_tested = if_else(hivtstever == 2, 1, 0, missing = 0),
            reported_negative = if_else(hivpsbp == 2 | (hivtsbp = 1 & hivrtpg == 2) | hivrslr == 2 | hivtstrslt == 2, 1, 0, missing = 0),
            pos_test_year = if_else(hivtfposy > 0, 1, 0),
            tested12mo = if_else(testedreceiveddetail %in% c(1,4), 1, 0),
            cd4testever = recode_one_zero(cd4testever),
            arvstakenev = if_else(arvstakenev == 1, 1, 0, missing = 0),
            arvscurrent = if_else(arvscurrent == 1, 1, 0, missing = 0),
            oth_art_care = if_else(arvtklb == 1 | arvcntn == 1 | hivcare == 1, 1, 0, missing = 0),
            across(-personid, recode_one_zero))

bio_select_var = bio %>% 
  filter(bt_status == 1,
         hivstatusfinal == 1) %>% 
  select(personid, awareartselfreported)
```

```{r}
modelvars <- c( "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "vlunder200")
# Remove all rows which have values other than 0 or 1
aware_patterns <- bio_select_mod %>% 
  select( all_of(modelvars) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
awareARTVL.lca3 <- randomLCA(aware_patterns, nclass = 3, quadpoints = 30)

outcomep <- round(awareARTVL.lca3$outcomep, 4)
colnames(outcomep) <- names(aware_patterns)

OutcomeProbs.Table.AwareARTVL <- as.data.frame( cbind( Class = c(1, 2, 3), outcomep)) %>%
  mutate(across(-Class, as.numeric))
```

```{r}
modelvars_arv <- c( "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "arvstatus")
# Remove all rows which have values other than 0 or 1
aware_patterns_arv <- bio_select_mod %>% 
  select( all_of(modelvars_arv) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
awareART.lca3 <- randomLCA(aware_patterns_arv, nclass = 3, quadpoints = 30)

outcomep <- round(awareART.lca3$outcomep, 4)
colnames(outcomep) <- names(aware_patterns_arv)

OutcomeProbs.Table.AwareART <- as.data.frame( cbind( Class = c(1, 2, 3), outcomep)) %>%
  mutate(across(-Class, as.numeric))
```

```{r}
set.seed(719)
lca_out_full = postClassProbs(awareARTVL.lca3) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3, 1,
                                 if_else(class_3 > class_1, 3, 1)), 
                         if_else(class_2 > class_3, 2, 3)))
```

```{r}
set.seed(719)
lca_out_full_arv = postClassProbs(awareART.lca3) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3, 1,
                                 if_else(class_3 > class_1, 3, 1)), 
                         if_else(class_2 > class_3, 2, 3)))
```

```{r}
full_dat_out = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_full, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "vlunder200"))
```

```{r}
full_dat_out_arv = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_full_arv, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "arvstatus"))
```

```{r}
class_out = full_join(full_dat_out, full_dat_out_arv, by = c("personid", "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "arvstatus", "vlunder200", "awareartselfreported.y"))
```

```{r}
modelvars_simple <- c( "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care")
# Remove all rows which have values other than 0 or 1
aware_patterns_simple <- bio_select_mod %>% 
  select( all_of(modelvars_simple) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
aware.lca_simple <- randomLCA(aware_patterns_simple, nclass = 3, quadpoints = 30)

outcomep <- round(aware.lca_simple$outcomep, 4)
colnames(outcomep) <- names(aware_patterns_simple)

OutcomeProbs.Table.Aware <- as.data.frame( cbind( Class = c(1, 2, 3), outcomep)) %>%
  mutate(across(-Class, as.numeric))
```

```{r}
set.seed(719)
lca_out_simple = postClassProbs(aware.lca_simple) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3, 1,
                                 if_else(class_3 > class_1, 3, 1)), 
                         if_else(class_2 > class_3, 2, 3)))
```

```{r}
simple_dat_out = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_simple, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care"))
```

```{r}
simple_vls_class_out = full_join(full_dat_out, simple_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "arvstatus", "awareartselfreported.y"))

simple_arv_class_out = full_join(full_dat_out_arv, simple_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "awareartselfreported.y"))
```

```{r}
gt(OutcomeProbs.Table.AwareARTVL)
gt(OutcomeProbs.Table.AwareART)
gt(OutcomeProbs.Table.Aware)
```

# compare different models 

Cross tabulate to compare different models

### vlunder200 model vs. no lab data model

```{r}
xtabs(~class.x + class.y, data = simple_vls_class_out)
```

### arvstatus model vs. no lab data

```{r}
xtabs(~class.x + class.y, data = simple_arv_class_out)
```

## awareartselfreported

no lab data

```{r}
xtabs(~awareartselfreported.y + class, data = simple_dat_out, exclude = "99")
```

vlunder200 model

```{r}
xtabs(~awareartselfreported.y + class, data = full_dat_out, exclude = "99")
```

arvstatus model

```{r}
xtabs(~awareartselfreported.y + class, data = full_dat_out_arv, exclude = "99")
```

## compare models to tri90 variables

### tri90aware

arvstatus model

```{r}
conmat_aware_df_arv = full_dat_out_arv

# 1 - unaware
# 2 - aware not on art
# 3 - aware on art

conmat_aware_df_arv$class = factor(conmat_aware_df_arv$class, levels = c("unaware" = 1, "aware - not on art" = 2, "aware - on art" = 3), labels = c("unaware", "aware", "aware"))
conmat_aware_df_arv$tri90aware = factor(conmat_aware_df_arv$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(conmat_aware_df_arv$class, conmat_aware_df_arv$tri90aware, positive = "aware")
cMat
```

vlunder200 model

```{r}
conmat_aware_df = full_dat_out

# 1 - aware not on art
# 2 - aware on art
# 3 - unaware

conmat_aware_df$class = factor(conmat_aware_df$class, levels = c("aware - not on art" = 1, "aware - on art" = 2, "unaware" = 3), labels = c("aware", "aware", "unaware"))
conmat_aware_df$tri90aware = factor(conmat_aware_df$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(conmat_aware_df$class, conmat_aware_df$tri90aware, positive = "aware")
cMat
```

no lab data model

```{r}
simple_conmat_df = simple_dat_out

# 1 - aware on art
# 2- aware not on art
# 3 - unaware

simple_conmat_df$class = factor(simple_conmat_df$class, levels = c("aware - on art" = 1, "aware - not on art" = 2, "unaware" = 3), labels = c("aware", "aware", "unaware"))
simple_conmat_df$tri90aware = factor(simple_conmat_df$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(simple_conmat_df$class, simple_conmat_df$tri90aware, positive = "aware")
cMat

```

## tri90art

arvstatus model

```{r}
conmat_aware_df_arv = full_dat_out_arv

# 1 - unaware
# 2 - aware not on art
# 3 - aware on art

conmat_aware_df_arv$class = factor(conmat_aware_df_arv$class, levels = c("unaware" = 1, "aware - not on art" = 2, "aware - on art" = 3), labels = c("not on art", "not on art", "on art"))
conmat_aware_df_arv$tri90art = factor(conmat_aware_df_arv$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(conmat_aware_df_arv$class, conmat_aware_df_arv$tri90art, positive = "on art")
cMat
```

vlunder200 model

```{r}
conmat_aware_df = full_dat_out

# 1 - aware not on art
# 2 - aware on art
# 3 - unaware

conmat_aware_df$class = factor(conmat_aware_df$class, levels = c("aware - not on art" = 1, "aware - on art" = 2, "unaware" = 3), labels = c("not on art", "on art", "not on art"))
conmat_aware_df$tri90art = factor(conmat_aware_df$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(conmat_aware_df$class, conmat_aware_df$tri90art, positive = "on art")
cMat
```

no lab data model

```{r}
simple_conmat_df = simple_dat_out

# 1 - aware on art
# 2- aware not on art
# 3 - unaware

simple_conmat_df$class = factor(simple_conmat_df$class, levels = c("aware - on art" = 1, "aware - not on art" = 2, "unaware" = 3), labels = c("on art", "not on art", "not on art"))
simple_conmat_df$tri90art = factor(simple_conmat_df$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(simple_conmat_df$class, simple_conmat_df$tri90art, positive = "on art")
cMat

```

# simpler model

Remove "arvstakenev" and "oth_art_care" variables.

## LCA, simpler

### vlunder200

```{r}
modelvars_simpler_vl <- c( "never_tested", "reported_negative", "reported_positive", "arvscurrent", "vlunder200")
# Remove all rows which have values other than 0 or 1
aware_patterns_simpler_vl <- bio_select_mod %>% 
  select( all_of(modelvars_simpler_vl) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
aware.lca_simpler_vl <- randomLCA(aware_patterns_simpler_vl, nclass = 3, quadpoints = 30)

outcomep <- round(aware.lca_simpler_vl$outcomep, 4)
colnames(outcomep) <- names(aware_patterns_simpler_vl)

OutcomeProbs.Table.Aware_simpler_vl <- as.data.frame( cbind( Class = c(1, 2, 3), outcomep)) %>%
  mutate(across(-Class, as.numeric))
```

### arvstatus (3 classes, 5 variables)

```{r}
modelvars_simpler_arv3 <- c( "never_tested", "reported_negative", "reported_positive", "arvscurrent", "arvstatus")
# Remove all rows which have values other than 0 or 1
aware_patterns_simpler_arv3 <- bio_select_mod %>% 
  select( all_of(modelvars_simpler_arv3) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
aware.lca_simpler_arv3 <- randomLCA(aware_patterns_simpler_arv3, nclass = 3, quadpoints = 30)

outcomep_simpler_arv3 <- round(aware.lca_simpler_arv3$outcomep, 4)
colnames(outcomep_simpler_arv3) <- names(aware_patterns_simpler_arv3)

OutcomeProbs.Table.Aware_simpler_arv3 <- as.data.frame( cbind( Class = c(1, 2, 3), outcomep_simpler_arv3)) %>%
  mutate(across(-Class, as.numeric))
```

### arvstatus (4 classes)

```{r}
modelvars_simpler_arv <- c( "never_tested", "reported_negative", "reported_positive", "arvscurrent", "arvstatus")
# Remove all rows which have values other than 0 or 1
aware_patterns_simpler_arv <- bio_select_mod %>% 
  select( all_of(modelvars_simpler_arv) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
aware.lca_simpler_arv <- randomLCA(aware_patterns_simpler_arv, nclass = 4, quadpoints = 30)

outcomep_simpler_arv <- round(aware.lca_simpler_arv$outcomep, 4)
colnames(outcomep_simpler_arv) <- names(aware_patterns_simpler_arv)

OutcomeProbs.Table.Aware_simpler_arv <- as.data.frame( cbind( Class = c(1, 2, 3, 4), outcomep_simpler_arv)) %>%
  mutate(across(-Class, as.numeric))
```


## extract classes

```{r}
set.seed(719)
lca_out_simpler_vl = postClassProbs(aware.lca_simpler_vl) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3, 1,
                                 if_else(class_3 > class_1, 3, 1)), 
                         if_else(class_2 > class_3, 2, 3)))
         # vlsaware = if_else(class == 1, 1, 0), # coded "forwards" meaning 1 is aware on art, 0 is EITHER unaware OR "aware and not on art"
         # awarenotart = if_else(class == 2, 0, 1), # coded "backwards" meaning 0 is "aware not on art", 1 is not "aware not on art", or in other words, "1" is NEITHER unaware NOR aware and not on art - in terms of  class, a "1" means that an individual may be EITHER "aware AND on art" OR unaware
         # unaware = if_else(class == 3, 0, 1)) # coded "backwards" meaning 0 is unaware, 1 is not unaware
```

### What are the classes in the vlunder200 model with fewer variables?

They don't seem to make much sense. Perhaps the first one is never tested, reported positive, not on arvs. The second one is reported negative, not on arvs, and the third one is reported positive, on arvs.

Maybe 2 classes would work better...

```{r}
gt(OutcomeProbs.Table.Aware_simpler_vl)
```

## fewer variable model with 4 classes

It appears 4 classes seems to do the trick...class 1 - never tested, class 2 - reported negative, class 3, aware not on art, class 4, aware on art.

### vlunder200 (5 variable)

```{r}
modelvars_simpler_vl <- c( "never_tested", "reported_negative", "reported_positive", "arvscurrent", "vlunder200")
# Remove all rows which have values other than 0 or 1
aware_patterns_simpler_vl <- bio_select_mod %>% 
  select( all_of(modelvars_simpler_vl) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
aware.lca_simpler_vl2 <- randomLCA(aware_patterns_simpler_vl, nclass = 4, quadpoints = 30)

outcomep_svl2 <- round(aware.lca_simpler_vl2$outcomep, 4)
colnames(outcomep_svl2) <- names(aware_patterns_simpler_vl)

OutcomeProbs.Table.Aware_simpler_vl2 <- as.data.frame( cbind( Class = c(1, 2, 3, 4), outcomep_svl2)) %>%
  mutate(across(-Class, as.numeric))
gt(OutcomeProbs.Table.Aware_simpler_vl2)
```

```{r}
postClassProbs(aware.lca_simpler_vl2) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3,
                                 if_else(class_3 > class_4, 1,
                                         if_else(class_4 > class_1, 4, 1)),
                                 if_else(class_3 > class_1,
                                         if_else(class_1 > class_4, 3,
                                                 if_else(class_4 > class_3, 4, 3)),
                                         if_else(class_1 > class_4, 1, 4))),
                         if_else(class_2 > class_3, 
                                 if_else(class_3 > class_4, 2,
                                         if_else(class_4 > class_2, 4, 2)),
                                 if_else(class_4 > class_3, 4, 3)))) %>% 
  select(never_tested:freq, class) %>% 
  arrange(desc(vlunder200)) %>% 
  gt() %>% 
  tab_row_group(group = "class_1",
                rows = c(6, 7, 13)) %>% 
  tab_row_group(group = "class_2",
                rows = c(2, 5)) %>% 
  tab_row_group(group = "class_3",
                rows = c(3, 4, 10)) %>% 
  tab_row_group(group = "class_4",
                rows = c(1, 8, 9, 11, 12)) %>% 
  tab_spanner(label = "profile",
              columns = vars(never_tested, reported_negative, reported_positive, arvscurrent, vlunder200)) %>% 
  tab_spanner(label = "cases",
              columns = vars(freq)) %>% 
  tab_spanner(label = "LCA Class",
              columns = vars(class))
```


## worth a try to add another class?

Maybe compare the models using 4 classes so they maintain some of the original intended class structure - to compare tri90aware and tri90art variables.

For the vlunder200 full model classes would apparently be: 1 - aware, not on art; 2 - reported negative; 3 - never tested; 4 - aware on art.

```{r}
modelvars <- c( "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "vlunder200")
# Remove all rows which have values other than 0 or 1
aware_patterns <- bio_select_mod %>% 
  select( all_of(modelvars) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
awareARTVL.lca4 <- randomLCA(aware_patterns, nclass = 4, quadpoints = 30)

outcomep4 <- round(awareARTVL.lca4$outcomep, 4)
colnames(outcomep4) <- names(aware_patterns)

OutcomeProbs.Table.AwareARTVL4 <- as.data.frame( cbind( Class = c(1, 2, 3, 4), outcomep4)) %>%
  mutate(across(-Class, as.numeric))
gt(OutcomeProbs.Table.AwareARTVL4)
```

For the arvstatus-model, classes could be viewed as: 1 - never tested; 2 - aware on art; 3 - aware not on art; 4 - reported negative

This might actually work...?

```{r}
modelvars <- c( "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "arvstatus")
# Remove all rows which have values other than 0 or 1
aware_patterns_arv4 <- bio_select_mod %>% 
  select( all_of(modelvars) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
awareART.lca4 <- randomLCA(aware_patterns_arv4, nclass = 4, quadpoints = 30)

outcomep_arv4 <- round(awareART.lca4$outcomep, 4)
colnames(outcomep_arv4) <- names(aware_patterns_arv4)

OutcomeProbs.Table.AwareART4 <- as.data.frame( cbind( Class = c(1, 2, 3, 4), outcomep_arv4)) %>%
  mutate(across(-Class, as.numeric))
gt(OutcomeProbs.Table.AwareART4)
```

For the simpler model (no lab work), the classes are: 1- aware on art; 2 - reported negative; 3 - aware not on art; and 4 - never tested.

Okay, so let's compare them using 4 classes.

```{r}
modelvars_simple <- c( "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care")
# Remove all rows which have values other than 0 or 1
aware_patterns_simple4 <- bio_select_mod %>% 
  select( all_of(modelvars_simple) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
aware.lca_simple4 <- randomLCA(aware_patterns_simple, nclass = 4, quadpoints = 30)

outcomep_simple4 <- round(aware.lca_simple4$outcomep, 4)
colnames(outcomep_simple4) <- names(aware_patterns_simple4)

OutcomeProbs.Table.Aware4 <- as.data.frame( cbind( Class = c(1, 2, 3, 4), outcomep_simple4)) %>%
  mutate(across(-Class, as.numeric))
gt(OutcomeProbs.Table.Aware4)
```

## extract classes from the models with 4 LCA classes

```{r}
# simpler_vl2 - 4 variables + lab data
set.seed(719)
lca_out_simpler_vl2 = postClassProbs(aware.lca_simpler_vl2) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3,
                                 if_else(class_3 > class_4, 1,
                                         if_else(class_4 > class_1, 4, 1)),
                                 if_else(class_3 > class_1,
                                         if_else(class_1 > class_4, 3,
                                                 if_else(class_4 > class_3, 4, 3)),
                                         if_else(class_1 > class_4, 1, 4))),
                         if_else(class_2 > class_3, 
                                 if_else(class_3 > class_4, 2,
                                         if_else(class_4 > class_2, 4, 2)),
                                 if_else(class_4 > class_3, 4, 3))))

# simple - 5 SR variables, no lab data
set.seed(719)
lca_out_simple4 = postClassProbs(aware.lca_simple4) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3,
                                 if_else(class_3 > class_4, 1,
                                         if_else(class_4 > class_1, 4, 1)),
                                 if_else(class_3 > class_1,
                                         if_else(class_1 > class_4, 3,
                                                 if_else(class_4 > class_3, 4, 3)),
                                         if_else(class_1 > class_4, 1, 4))),
                         if_else(class_2 > class_3, 
                                 if_else(class_3 > class_4, 2,
                                         if_else(class_4 > class_2, 4, 2)),
                                 if_else(class_4 > class_3, 4, 3))))

# vlunder200 - full model (7 variables), vlunder200
set.seed(719)
lca_out_vls4 = postClassProbs(awareARTVL.lca4) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3,
                                 if_else(class_3 > class_4, 1,
                                         if_else(class_4 > class_1, 4, 1)),
                                 if_else(class_3 > class_1,
                                         if_else(class_1 > class_4, 3,
                                                 if_else(class_4 > class_3, 4, 3)),
                                         if_else(class_1 > class_4, 1, 4))),
                         if_else(class_2 > class_3, 
                                 if_else(class_3 > class_4, 2,
                                         if_else(class_4 > class_2, 4, 2)),
                                 if_else(class_4 > class_3, 4, 3))))

# arvstatus - full model (7 variables), arvstatus
set.seed(719)
lca_out_art4 = postClassProbs(awareART.lca4) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3,
                                 if_else(class_3 > class_4, 1,
                                         if_else(class_4 > class_1, 4, 1)),
                                 if_else(class_3 > class_1,
                                         if_else(class_1 > class_4, 3,
                                                 if_else(class_4 > class_3, 4, 3)),
                                         if_else(class_1 > class_4, 1, 4))),
                         if_else(class_2 > class_3, 
                                 if_else(class_3 > class_4, 2,
                                         if_else(class_4 > class_2, 4, 2)),
                                 if_else(class_4 > class_3, 4, 3))))

# simpler model (5 variables), arvstatus
set.seed(719)
lca_out_simpler_arv = postClassProbs(aware.lca_simpler_arv) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3,
                                 if_else(class_3 > class_4, 1,
                                         if_else(class_4 > class_1, 4, 1)),
                                 if_else(class_3 > class_1,
                                         if_else(class_1 > class_4, 3,
                                                 if_else(class_4 > class_3, 4, 3)),
                                         if_else(class_1 > class_4, 1, 4))),
                         if_else(class_2 > class_3, 
                                 if_else(class_3 > class_4, 2,
                                         if_else(class_4 > class_2, 4, 2)),
                                 if_else(class_4 > class_3, 4, 3))))
```

# 4 LCA classes

## compare to no-lab-data-, vlunder200-, arvstatus-models

vlunder200 (5 variable)

```{r}
simpler_vl2_dat_out = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_simpler_vl2, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvscurrent", "vlunder200"))
```

vlunder200 (7 variable)

```{r}
full_dat_out4 = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_vls4, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "vlunder200"))

```

arvstatus(7 variable)

```{r}
full_dat_out_arv4 = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_art4, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "arvstatus"))
```

no lab data

```{r}
simple_out4 = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_simple4, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care"))
```

```{r, echo = FALSE}
simpler_vl2_vls_class_out = full_join(full_dat_out4, simpler_vl2_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "arvscurrent", "arvstatus", "awareartselfreported.y"))

simpler_vl2_arv_class_out = full_join(full_dat_out_arv4, simpler_vl2_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "vlunder200", "arvscurrent", "awareartselfreported.y"))

simpler_vl2_simple_class_out = full_join(simple_out4, simpler_vl2_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "vlunder200", "arvscurrent", "arvstatus", "awareartselfreported.y"))

vls_arv_class_out = full_join(full_dat_out4, full_dat_out_arv4, by = c("personid", "never_tested", "reported_negative", "reported_positive", "vlunder200", "arvstatus", "arvscurrent", "awareartselfreported.y"))
```

## cross tabulation

Let's check the full datasets first out of curiosity.

Cross tabulate the 7 variable models now using 4 classes.

There is similar agreement between classes to the 3 class model.

```{r, echo = TRUE}
xtabs(~class.x + class.y, data = vls_arv_class_out)
```

What about other models? (comparing simplified models - no lab data, or fewer variables)

There are more disagreements between classes. But, how does it compare to tri90 variables?

vlunder200 (5 variable) vs. vlunder200 (7 variable)

```{r, echo = TRUE}
xtabs(~class.x + class.y, data = simpler_vl2_vls_class_out)
```

vlunder200 (5 variable) vs. arvstatus(7 variable)

```{r, echo = TRUE}
xtabs(~class.x + class.y, data = simpler_vl2_arv_class_out)
```

vlunder200 (5 variable) vs. no lab data

```{r, echo = TRUE}
xtabs(~class.x + class.y, data = simpler_vl2_simple_class_out)
```

## Reminder of class labels for each LCA output

```{r, echo = TRUE}
gt(OutcomeProbs.Table.Aware4)
gt(OutcomeProbs.Table.AwareART4)
gt(OutcomeProbs.Table.AwareARTVL4)
gt(OutcomeProbs.Table.Aware_simpler_vl2)
```

# tri90variables

## vlunder200 (5 variables)

tri90aware

```{r, echo = TRUE}
simpler_vl2_conmat_df = simpler_vl2_dat_out

# 1 - never tested
# 2 - aware on art
# 3 - reported negative
# 4 - aware not on art

simpler_vl2_conmat_df$class = factor(simpler_vl2_conmat_df$class, levels = c("never tested" = 1, "aware on art" = 2, "reported negative" = 3, "aware not on art" = 4), labels = c("unaware", "aware", "unaware", "aware"))
simpler_vl2_conmat_df$tri90aware = factor(simpler_vl2_conmat_df$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(simpler_vl2_conmat_df$class, simpler_vl2_conmat_df$tri90aware, positive = "aware")
cMat
```

tri90art

```{r, echo = TRUE}
simpler_vl2_conmat_df = simpler_vl2_dat_out

# 1 - never tested
# 2 - aware on art
# 3 - reported negative
# 4 - aware not on art

simpler_vl2_conmat_df$class = factor(simpler_vl2_conmat_df$class, levels = c("never tested" = 1, "aware on art" = 2, "reported negative" = 3, "aware not on art" = 4), labels = c("not on art", "on art", "not on art", "not on art"))
simpler_vl2_conmat_df$tri90art = factor(simpler_vl2_conmat_df$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(simpler_vl2_conmat_df$class, simpler_vl2_conmat_df$tri90art, positive = "on art")
cMat
```

## no lab data model

tri90art

```{r, echo = TRUE}
simple_out4_conmat_df = simple_out4

# 1 - aware on art
# 2 - never tested
# 3 - aware not on art
# 4 = reported negative

simple_out4_conmat_df$class = factor(simple_out4_conmat_df$class, levels = c("aware on art" = 1, "never tested" = 2, "aware not on art" = 3, "reported negative" = 4), labels = c("on art", "not on art", "not on art", "not on art"))
simple_out4_conmat_df$tri90art = factor(simple_out4_conmat_df$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(simple_out4_conmat_df$class, simple_out4_conmat_df$tri90art, positive = "on art")
cMat
```

tri90aware

```{r, echo = TRUE}
simple_out4_conmat_df = simple_out4

# 1 - aware on art
# 2 - never tested
# 3 - aware not on art
# 4 = reported negative

simple_out4_conmat_df$class = factor(simple_out4_conmat_df$class, levels = c("aware on art" = 1, "never tested" = 2, "aware not on art" = 3, "reported negative" = 4), labels = c("aware", "unaware", "aware", "unaware"))
simple_out4_conmat_df$tri90aware = factor(simple_out4_conmat_df$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(simple_out4_conmat_df$class, simple_out4_conmat_df$tri90aware, positive = "aware")
cMat
```

## vlunder200 full model (7 variables)

tri90aware

```{r, echo = TRUE}
full_dat_out4_conmat_df = full_dat_out4

# 1 - reported negative
# 2 - aware on art
# 3 - aware not on art
# 4 - tested negative

full_dat_out4_conmat_df$class = factor(full_dat_out4_conmat_df$class, levels = c("reported negative" = 1, "aware on art" = 2, "aware not on art" = 3, "tested negative" = 4), labels = c("unaware", "aware", "aware", "unaware"))
full_dat_out4_conmat_df$tri90aware = factor(full_dat_out4_conmat_df$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(full_dat_out4_conmat_df$class, full_dat_out4_conmat_df$tri90aware, positive = "aware")
cMat
```

tri90art

```{r, echo = TRUE}
full_dat_out4_conmat_df = full_dat_out4

# 1 - reported negative
# 2 - aware on art
# 3 - aware not on art
# 4 - tested negative

full_dat_out4_conmat_df$class = factor(full_dat_out4_conmat_df$class, levels = c("reported negative" = 1, "aware on art" = 2, "aware not on art" = 3, "tested negative" = 4), labels = c("not on art", "on art", "not on art", "not on art"))
full_dat_out4_conmat_df$tri90art = factor(full_dat_out4_conmat_df$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(full_dat_out4_conmat_df$class, full_dat_out4_conmat_df$tri90art, positive = "on art")
cMat
```

## arvstatus model (7 variables)

tri90aware

```{r, echo = TRUE}
full_dat_arv4_conmat_df = full_dat_out_arv4

# 1 - aware on art
# 2 - reported negative
# 3 - aware not on art
# 4 = never tested

full_dat_arv4_conmat_df$class = factor(full_dat_arv4_conmat_df$class, levels = c("aware on art" = 1, "reported negative" = 2, "aware not on art" = 3, "never tested" = 4), labels = c("aware", "unaware", "aware", "unaware"))
full_dat_arv4_conmat_df$tri90aware = factor(full_dat_arv4_conmat_df$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(full_dat_arv4_conmat_df$class, full_dat_arv4_conmat_df$tri90aware, positive = "aware")
cMat
```

tri90art

```{r, echo = TRUE}
full_dat_arv4_conmat_df = full_dat_out_arv4

# 1 - aware on art
# 2 - reported negative
# 3 - aware not on art
# 4 = never tested

full_dat_arv4_conmat_df$class = factor(full_dat_arv4_conmat_df$class, levels = c("aware on art" = 1, "reported negative" = 2, "aware not on art" = 3, "never tested" = 4), labels = c("on art", "not on art", "not on art", "not on art"))
full_dat_arv4_conmat_df$tri90art = factor(full_dat_arv4_conmat_df$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(full_dat_arv4_conmat_df$class, full_dat_arv4_conmat_df$tri90art, positive = "on art")
cMat
```

# build the arvstatus model (5 variables)

```{r}
simpler_arv_dat_out = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_simpler_arv, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvscurrent", "arvstatus"))
```


```{r}
simpler_arv_full_class_out = full_join(full_dat_out4, simpler_arv_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "arvscurrent", "arvstatus", "awareartselfreported.y"))

simpler_arv_arv_class_out = full_join(full_dat_out_arv4, simpler_arv_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "vlunder200", "arvscurrent", "awareartselfreported.y"))

simpler_arv_simple_class_out = full_join(simple_out4, simpler_arv_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "vlunder200", "arvscurrent", "arvstatus", "awareartselfreported.y"))

simpler_vls_arv_class_out = full_join(simpler_vl2_dat_out, simpler_arv_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "vlunder200", "arvstatus", "arvscurrent", "awareartselfreported.y"))
```

## Compare the 5-variable models

vlunder200 (5 variables) vs. arvstatus (5 variables)

Some disagreement. How about the other models?

```{r, echo = TRUE}
xtabs(~class.x + class.y, data = simpler_vls_arv_class_out)
```

arvstatus (5 variable) vs. vlunder (7 variable)

```{r ,echo = TRUE}
xtabs(~class.x + class.y, data = simpler_arv_full_class_out)
```

arvstatus (5 variable) vs. arvstatus (7 variable)

```{r ,echo = TRUE}
xtabs(~class.x + class.y, data = simpler_arv_arv_class_out)
```

arvstatus (5 variable) vs. no lab data

```{r ,echo = TRUE}
xtabs(~class.x + class.y, data = simpler_arv_simple_class_out)
```

# arvstatus (5 variable)



reminder of classes

```{r, echo = TRUE}
gt(OutcomeProbs.Table.Aware_simpler_arv)
```

## compare to tri90 variables

### tri90aware

#### cross-tabulation

tri90art

```{r, echo = TRUE}
simpler_arv_conmat_df = simpler_arv_dat_out

# 1 - reported negative
# 2 - never tested
# 3 - aware not on art
# 4 - aware on art

simpler_arv_conmat_df$class = factor(simpler_arv_conmat_df$class, levels = c("reported negative" = 1, "never tested" = 2, "aware not on art" = 3, "aware on art" = 4), labels = c("not on art", "not on art", "not on art", "on art"))
simpler_arv_conmat_df$tri90art = factor(simpler_arv_conmat_df$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(simpler_arv_conmat_df$class, simpler_arv_conmat_df$tri90art, positive = "on art")
cMat
```

tri90aware

```{r, echo = TRUE}
simpler_arv_conmat_df = simpler_arv_dat_out

# 1 - reported negative
# 2 - never tested
# 3 - aware not on art
# 4 - aware on art

simpler_arv_conmat_df$class = factor(simpler_arv_conmat_df$class, levels = c("reported negative" = 1, "never tested" = 2, "aware not on art" = 3, "aware on art" = 4), labels = c("unaware", "unaware", "aware", "aware"))
simpler_arv_conmat_df$tri90aware = factor(simpler_arv_conmat_df$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(simpler_arv_conmat_df$class, simpler_arv_conmat_df$tri90aware, positive = "aware")
cMat
```

NOTE: The 5 variables models appear to not classify respondents as well possibly because there are two classes of aware individuals with high rates of arvstatus.

In Eswatini, it appears a 7 variable model works better.


# check BIC

```{r}
tibble("3_classes" = 
         c(BIC(aware.lca_simple),BIC(awareARTVL.lca3),BIC(awareART.lca3),BIC(aware.lca_simpler_vl),BIC(aware.lca_simpler_arv3)),
       "model_3" = list("lca_nolabdat","lca_vlunder200","lca_arvstatus","lca_simple_vlunder200","lca_simple_arvstatus"),
       "4_classes" = list(BIC(aware.lca_simple4),BIC(awareARTVL.lca4),BIC(awareART.lca4),BIC(aware.lca_simpler_vl2),BIC(aware.lca_simpler_arv)),
       "model_4" = list("lca_nolabdat","lca_vlunder200","lca_arvstatus","lca_simple_vlunder200", "lca_simple_arvstatus")) %>% 
  unnest()
```

```{r}
# Eswatini
## no lab data
# postClassProbs(aware.lca_simple4) %>% 
#   janitor::clean_names() %>% 
#   mutate(class = if_else(class_1 > class_2, 
#                          if_else(class_2 > class_3,
#                                  if_else(class_3 > class_4, 1,
#                                          if_else(class_4 > class_1, 4, 1)),
#                                  if_else(class_3 > class_1,
#                                          if_else(class_1 > class_4, 3,
#                                                  if_else(class_4 > class_3, 4, 3)),
#                                          if_else(class_1 > class_4, 1, 4))),
#                          if_else(class_2 > class_3, 
#                                  if_else(class_3 > class_4, 2,
#                                          if_else(class_4 > class_2, 4, 2)),
#                                  if_else(class_4 > class_3, 4, 3)))) %>% 
#   select(never_tested:freq, class)

## vlunder200 (7 variables)
postClassProbs(awareARTVL.lca4) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3,
                                 if_else(class_3 > class_4, 1,
                                         if_else(class_4 > class_1, 4, 1)),
                                 if_else(class_3 > class_1,
                                         if_else(class_1 > class_4, 3,
                                                 if_else(class_4 > class_3, 4, 3)),
                                         if_else(class_1 > class_4, 1, 4))),
                         if_else(class_2 > class_3, 
                                 if_else(class_3 > class_4, 2,
                                         if_else(class_4 > class_2, 4, 2)),
                                 if_else(class_4 > class_3, 4, 3)))) %>% 
  select(never_tested:freq, class) %>% 
  group_by(class, vlunder200) %>%
  summarize(freq = sum(freq))

postClassProbs(awareARTVL.lca4) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3,
                                 if_else(class_3 > class_4, 1,
                                         if_else(class_4 > class_1, 4, 1)),
                                 if_else(class_3 > class_1,
                                         if_else(class_1 > class_4, 3,
                                                 if_else(class_4 > class_3, 4, 3)),
                                         if_else(class_1 > class_4, 1, 4))),
                         if_else(class_2 > class_3, 
                                 if_else(class_3 > class_4, 2,
                                         if_else(class_4 > class_2, 4, 2)),
                                 if_else(class_4 > class_3, 4, 3)))) %>% 
  select(never_tested:freq, class) %>% 
  arrange(class, vlunder200) %>% 
  gt() %>% 
  tab_row_group(group = "class_1",
                rows = c(1:9)) %>% 
  tab_row_group(group = "class_2",
                rows = c(10:13)) %>% 
  tab_row_group(group = "class_3",
                rows = c(14:23)) %>% 
  tab_row_group(group = "class_4",
                rows = c(24:25)) %>% 
  tab_spanner(label = "profile",
              columns = vars(never_tested, reported_negative, reported_positive, arvstakenev, arvscurrent, oth_art_care, vlunder200)) %>% 
  tab_spanner(label = "cases",
              columns = vars(freq)) %>% 
  tab_spanner(label = "LCA Class",
              columns = vars(class)) %>% 
  gtsave("20201130-class_profiles_lesotho.html")

postClassProbs(awareARTVL.lca4) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3,
                                 if_else(class_3 > class_4, 1,
                                         if_else(class_4 > class_1, 4, 1)),
                                 if_else(class_3 > class_1,
                                         if_else(class_1 > class_4, 3,
                                                 if_else(class_4 > class_3, 4, 3)),
                                         if_else(class_1 > class_4, 1, 4))),
                         if_else(class_2 > class_3, 
                                 if_else(class_3 > class_4, 2,
                                         if_else(class_4 > class_2, 4, 2)),
                                 if_else(class_4 > class_3, 4, 3)))) %>% 
  select(never_tested:freq, class) %>% 
  mutate(class = if_else(class %in% c(1, 3), "aware", "unaware")) %>% 
  group_by(class, vlunder200) %>% 
  summarize(freq = sum(freq)) %>% 
  group_by(class) %>% 
  mutate(p_vls = freq/sum(freq))
```

```{r}

```

