---
title: "phia_lca_jsg2145"
author: "Jared Garfinkel"
date: "8/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(randomLCA)
library(pROC)
library(gridExtra)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
# Read public-release datasets
#Biomarker dataset
biomarker <- read_csv(file = "./data/Shims22016adultbio.csv",
                      col_names = TRUE, 
                      col_types = NULL,
                      locale = default_locale(), 
                      na = c("", "NA"), 
                      quoted_na = TRUE,
                      quote = "\"", 
                      comment = "", 
                      trim_ws = TRUE, 
                      skip = 0,
                      n_max = Inf, 
                      progress = show_progress(), 
                      skip_empty_rows = TRUE)

#Individual interview dataset
individual <- read_csv(file = "./data/Shims22016adultind.csv",
                       col_names = TRUE, 
                       col_types = NULL,
                       locale = default_locale(), 
                       na = c("", "NA"), 
                       quoted_na = TRUE,
                       quote = "\"", 
                       comment = "", 
                       trim_ws = TRUE, 
                       skip = 0,
                       n_max = Inf, 
                       progress = show_progress(), 
                       skip_empty_rows = TRUE)
```

```{r clean data}
ind_dat <-
  filter(individual, bt_status == 1) %>% #
  select(personid, 
         gender, 
         age, 
         hivstatusfinal, 
         known_hiv_status, 
         arvscurrent) %>% 
  mutate(gender = factor(gender, 
                         labels = c("1" = "male",
                                    "2" = "female")),
         across(.funs = as_factor))

bio_dat <- biomarker %>% 
  filter(hiv1statusfinalsurvey == 1) %>%
  select(personid, 
         hiv1statusfinalsurvey,
         aware,
         art,
         awareselfreported, 
         arvstatus, 
         artselfreported, 
         resultvlc, 
         vls, 
         tri90, 
         tri90aware, 
         tri90art, 
         tri90vls) %>%
  mutate(resultvlc = recode(resultvlc, 
                    "< LLOD" = "1",
                    "< LLOQ: 20" = "20",
                    "< LLOQ: 40" = "40",
                    "< LLOQ: 400" = "400",
                    "< LLOQ: 839" = "839",
                    "> ULOQ 10000000" = "10000000"),
         resultvlc = as.numeric(resultvlc),
         vlunder200 = if_else(resultvlc < 200, 1, 2),
         across(.funs = as_factor))
```

```{r codebook}
# aware:                1 - Aware or considered aware because ARVs detectable 
#                       2 - Unaware and ARVs not detectable, or unaware and ARV testing results missing 
#                       99 - Missing
#
# art:                  1 - ARVs detectable, self-reported on ART, or both ARVs detectable and self-reported on ART 
#                       2 - Unaware or aware, ARVs not detectable and self-reported not on ART, or aware, missing ARV testing data and self-reported not on ART 
#                       99 - Missing
# awareselfreported:    1 - Self-report aware of HIV + status 
#                       2 - Self-report not aware of HIV + status 
#                       99 - missing
# artselfreported:      1 - On ART 
#                       2 - Not on ART 
#                       99 - Missing
# awareartselfreported: 1 - Self-report as not previously diagnosed 
#                       2 - Self-report as previously diagnosed, not on ART 
#                       3 - Previously diagnosed, on ART 
#                       99 - Missing, including incomplete tri90 information
# arvstatus:            1 - ARV detected 
#                       2 - ARV not detected 
#                       99 - Missing
# resultvlc:            > ULOQ 10000000 - Upper limit of quantification 10000000
#                       < LLOD - less than lower limit of detection 
#                       < LLOQ: 839 - less than lower limit of quantification of 839 
#                       < LLOQ: 400 - less than lower limit of quantification of 400
#                       < LLOQ: 40 - less than lower limit of quantification of 40 
#                       < LLOQ: 20 - less than lower limit of quantification of 20
# hivselfreport:        1 - Self-reported positive 
#                       2 - Self-reported negative 
#                       3 - Self-reported never tested or never received test result 
#                       99 - Missing
# tri90art              1 - ARVs detectable, self-reported on ART, or both ARVs detectable and self-reported on ART 
#                       2 - ARVs not detectable and self-reported not on ART or missing ARV data and self-reported not on ART 
#                       3 - Recoded as not on ART (unaware and ARVs not detectable or unaware and ARVs missing) 
#                       99 - Incomplete Tri90 information
#
```

## First 90 (awareness)

### pre-processed data

```{r awareselfreported tables}


# This table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and had an art identified by a blood test
tbl1 = xtabs(~ awareselfreported + tri90art, data = bio_dat)

# This contingency table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and the status of a dummy variable that measures whether they have suppressed viral loads
tbl2 = xtabs(~ awareselfreported + arvstatus, data = bio_dat)

# This contingency table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and reported that they were taking ART.
tbl3 = xtabs(~ awareselfreported + artselfreported, data = bio_dat)

# This contingency table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and the pre-processed self-reported awareness of their HIV status
tbl4 = xtabs(~ tri90aware + awareselfreported, data = bio_dat)
```

```{r tri90aware tables}
# Grand total
nrow(bio_dat) # 3,055

# This table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and had an art identified by a blood test
tbl1_tri90 = xtabs(~ tri90aware + tri90art, data = bio_dat)

# This contingency table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and the status of a dummy variable that measures whether they have suppressed viral loads
tbl2_tri90 = xtabs(~ tri90aware + arvstatus, data = bio_dat)

# This contingency table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and reported that they were taking ART.
tbl3_tri90 = xtabs(~ tri90aware + artselfreported, data = bio_dat)

# This contingency table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and the pre-processed self-reported awareness of their HIV status
tbl4_tri90 = xtabs(~ tri90aware + awareselfreported, data = bio_dat)

# This table shows the respondents processed tri90aware status and their viral load suppression (under 200 copies/milliliter)
xtabs(~ tri90aware + vlunder200, data = bio_dat)
```

### tri90art

This combined chart shows the contingency table of awareselfreported vs. tri90art along with tri90aware vs. tri90art. While there are missing values in awareselfreported, there are none in tri90aware. The 57 missing values from tri90art were classified as missing for tri90aware data. The 60 respondents who were on art were reclassified as tri90aware.

```{r}
g1 <- tableGrob(tbl1)
g2 <- tableGrob(tbl1_tri90)

haligned <- gtable_combine(g1,g2, along=1)
valigned <- gtable_combine(g1,g2, along=2)
grid.arrange(haligned, valigned, ncol=2, top = "tri90art, awareselfreported vs. tri90aware")
```

### arvstatus

This chart shows contingency tables for awareselfreported alongside those for tri90aware with respect to arvstatus. As before, those with detectable arvs in their blood sample were reclassified as tri90aware regardless of selfreported aware status. Those missing arvstatus data appear to have been treated or classified differently. 3 respondents with missing arvstatus data were classified as tri90aware compared to 50 self-reported aware who were missing arvstatus data. Meanwhile, there was 1 respondent missing arvstatus classified as not tri90aware compared to 6 of those who reported that they were unaware and had missing arvstatus data. Although 52 respondents with missing arvstatus data were considered null under tri90aware, none who were missing awareselfreported data were also missing arvstatus data. Finally, it appears that 6 respondents who had no detectable arvs were considered missing tri90aware data, even though only 1 who had no detectable arvs were missing awareselfreported data. So, overall the biggest change was between those missing arvstatus data who said they were aware, most of whom it appears were reclassified as missing data in tri90aware. 50/3055 = 1.6%. This would effectively reduce the prevalence of awareness by removing about 50 individuals from the analysis.

#### Why were subjects removed from the analysis?

Generally an observation was removed if they had incomplete tri90 data, including blood-test data. Specifically, observations to be excluded in the 90 90 90 analysis with ARV data due to incomplete data defined as: aware missing and ARV data missing, aware and both ARV and self-reported ART data missing, or ARVs detectable / self-reported on ART and VLS missing.

```{r}
h1 <- tableGrob(tbl2)
h2 <- tableGrob(tbl2_tri90)

haligned <- gtable_combine(h1,h2, along=1)
valigned <- gtable_combine(h1,h2, along=2)
grid.arrange(haligned, valigned, ncol=2, top = "arvstatus")
```

## artselfreported

44 respondents were removed from the aware group if they reported that they were aware and on art. 3 respondents were removed from the aware group if they were awareselfreported and had no detectable arvs. 64 respondents who reported that they were unaware were reclassified. Of those reclassified, it appears that 55 of them were classified as aware and 9 were classified as missing data and not included in the tri90 analysis. So, as about 50 respondents were reclassified as aware from artselfreported data, almost 50 respondents were removed from the tri90 analysis who had previously been considered aware. This is a differential misclassification and deserves further study.

```{r}
j1 <- tableGrob(tbl3)
j2 <- tableGrob(tbl3_tri90)

haligned <- gtable_combine(j1,j2, along=1)
valigned <- gtable_combine(j1,j2, along=2)
grid.arrange(haligned, valigned, ncol=2, top = "artselfreported")
```

## awareselfreported

```{r, include = FALSE, eval = FALSE}
k1 <- tableGrob(tbl4)
k2 <- tableGrob(tbl4_tri90)

haligned <- gtable_combine(k1,k2, along=1)
valigned <- gtable_combine(k1,k2, along=2)
grid.arrange(haligned, valigned, ncol=2, top = "awareselfreported")
```


### tri90 (processed) data

```{r contingency table data}
bio_cong_dat = biomarker %>% 
  filter(tri90 == 1) %>% 
  select(personid, 
         hiv1statusfinalsurvey,
         aware,
         art,
         awareselfreported, 
         arvstatus, 
         artselfreported, 
         resultvlc, 
         vls, 
         tri90, 
         tri90aware, 
         tri90art, 
         tri90vls) %>%
  mutate(resultvlc = recode(resultvlc, 
                    "< LLOD" = "1",
                    "< LLOQ: 20" = "20",
                    "< LLOQ: 40" = "40",
                    "< LLOQ: 400" = "400",
                    "< LLOQ: 839" = "839",
                    "> ULOQ 10000000" = "10000000"),
         resultvlc = as.numeric(resultvlc),
         vlunder200 = if_else(resultvlc < 200, 1, 2),
         across(.funs = as_factor))
```

```{r}
# Grand total
nrow(bio_cong_dat)

# while all the respondents classified as unaware under tri90aware were unaware under awareselfreported, 59 respondents who self-reported they were unaware were reclassified as aware, and 1 respondent classified as missing data was reclassified as aware under tri90aware. A total of 60 respondents were reclassified from their self-reported aware status.
xtabs(~tri90aware + awareselfreported, bio_cong_dat)

# While no respondents classified as unaware under tri90aware were asked for artselfreported data, 306 of those who reported they were not on art were reclassified as aware and 64 respondents with missing artselfreported data were reclassified as aware
xtabs(~tri90aware + artselfreported, bio_cong_dat)

# While no respondents who tested positive for art were classified as aware, 282 respondents who tested negative for art were reclassified as aware, and 346 respondents who tested negative for art were classified as unaware.
xtabs(~ tri90aware + tri90art, data = bio_cong_dat)

# There were 546 respondents with viral load above 200 who were classified as aware. This may be indicative of treatment non-compliance. There were 96 respondents with viral load under 200 classified as unaware. This may indicate the presence of elite suppressors or transient viral load suppression among respondents.
xtabs(~tri90aware + vlunder200, bio_cong_dat)

# There were 369 respondents with no detectable arvs reclassified as tri90aware and 3 respondents with missing arvstatus data classified as aware. Meanwhile, 1 respondent with missing arvstatus data was classified as unaware. These 4 respondents missing arvstatus data may or may not have been reclassified.
xtabs(~tri90aware + arvstatus, bio_cong_dat)

# This table shows that 444 respondents with viral load above 1000 copies/ milliliter were reclassified as aware, while only 35 individuals with viral load under 1000 copies per milliliter were classified as unaware.
xtabs(~tri90aware + vls, bio_cong_dat)

# These variables are in agreement
xtabs(~tri90aware + aware, data = bio_cong_dat)

```

# Differential misclassification

Differential misclassification occurs when misclassification of exposure is not equal between subjects that have or do not have the health outcome, or when misclassification of the health outcome is not equal between exposed and unexposed subjects.

Assume the tri90 classifications represent the truth. It is possible to compare the classifications of respondents using pre-processed measurements (i.e. of blood samples and survey responses) to the "true" values and assess whether the misclassification is differential or non-differential.

```{r}
nrow(bio_cong_dat) # 2,997

# The variable aware is useful here because it mirrors tri90aware, yet it shows that reclassification took place amont those self reported that they were unaware or were missing awareselfreported data. In fact, 59 respondents were reclassified from self-reportedly unaware to aware. None were reclassified from aware to unaware.
xtabs(~tri90aware + awareselfreported + aware, bio_cong_dat) %>% 
  ftable(.)

# This table shows that 4 respondents missing artselfreported data were classified as tri90aware who reported that they were aware. These individuals were asked whether they were on art, but the data is missing.
xtabs(~tri90aware + awareselfreported + artselfreported, bio_cong_dat) %>% 
  ftable(.)

# Some respondents marked that they were not on art and had no detectable arvs in their blood. All of them were classified as aware. All respondents who responded that they were not aware of their status were recoded as not on art because they were not aware of their status. 
xtabs(~tri90aware + awareselfreported + tri90art, bio_cong_dat) %>% 
  ftable(.)

# Of those with viral loads above 200 copies/milliliter, 532 reported that they were aware and were subsequently coded as tri90aware; 13 reported that they were unaware, but were ultimately coded as aware; 330 responded that they were unaware and were coded as unaware. So, 13/986 or 1.5% of those with elevated viral loads were reclassified. On the other hand, of those with suppressed viral loads, 46 reported they were unaware and were subsequently reclassified; 16 reported that they were unaware and found to be unaware. So, of those who self-reported that they were unaware, 62/405 or 15% had viral load under 200 copies/milliliter, and 59/405 or 15% were reclassified as tri90aware. Finally, of those who self-reported unaware and who were viral load suppressed, 46/62 or 74% were reclassified as aware.
xtabs(~tri90aware + awareselfreported + vlunder200, bio_cong_dat) %>% 
  ftable(.)

# There was 1 respondent with missing arvstatus information who selfreported that they were unaware and classified as unaware. None of those with missing arvstatus data were reclassified from their self-reported awareness. The same can be said of those who had no detectable arvs in their blood. Only those with detectable arvs were reclassified from unaware to aware.
xtabs(~tri90aware + awareselfreported + arvstatus, bio_cong_dat) %>% 
  ftable(.)

# This table shows that there were reclassifications based on vls in each category. 48 respondents were reclassified from self reported unaware to tri90aware who were viral load suppressed (less than 1000 copies/milliliter), while 11 who were not viral load suppressed (less than 1000 copies/milliliter) were also reclassified as tri90aware from self-reported unaware. One respondent who was not viral load suppressed was missing self-reported awareness data, but classified as tri90aware. Of those who were viral load suppressed 48/2242 or 2% were reclassified. Interestingly of those who self-reported that they were unaware, 83/405 or 20% were viral load suppressed (less than 1000 copies/milliliter). Furthermore, of those who were self-reported unaware, 59/405 were ultimately reclassified as aware. Of those who self reported unaware and had suppressed viral loads, 48/83 or 58% were reclassified.
xtabs(~tri90aware + awareselfreported + vls, bio_cong_dat) %>% 
  ftable(.)

# This table shows that of 59 individuals who self reported unaware, 46 were viral load suppressed (under 200 copies per milliliter) with detectable arvs, and 13 were not viral load suppressed even though they had detectable arvs in their blood. Compare this to 48 viral load suppressed (under 1000 copies per milliliter) with detectable arvs and 11 not viral load suppressed even though they had detectable arvs. No respondents were classified as tri90aware if they did not have detectable arvs and reported that they were unaware. However, 59  respondents who said they were unaware were reclassified. Of those who reported that they were unaware, 59/405 or 15% were ultimately reclassified.
xtabs(~tri90aware + awareselfreported + vlunder200 + arvstatus, bio_cong_dat) %>% 
  ftable(., col.vars = c("vlunder200", "arvstatus"))

xtabs(~tri90aware + awareselfreported + vls + arvstatus, bio_cong_dat) %>% 
  ftable(., col.vars = c("vls", "arvstatus"))

# No respondents who self-reported that they were unaware were asked about art. So, respondents could only be reclassified if they self-reported that they were aware of their status. A respondent could not be reclassified based on self-reported art status if they reported that they were unaware of their hiv status. The variable artselfreported is only therefore capable of telling if someone has said that they are on art when they are on art. More specifically, the artselfreported variable will never reveal whether someone is saying something other than the truth about their awareness of their hiv status, but only what we find out through the arv detection test. If someone has self-reported that they are on arvs, then in every instance they were classified as tri90art. However, if they said they were self-reported not on art, then in 24/306 or 7.8% of cases, they were reclassified as tri90art. If they were missing self-reported art status then in 64/410 or 16% of cases they were classified as tri90art. Otherwise, they were classified as not on art due to missing self-reported data.
xtabs(~tri90aware + awareselfreported + artselfreported + tri90art, bio_cong_dat) %>% 
  ftable(., col.vars = c("artselfreported", "tri90art"))

# Of those self-reported on art, 90/2281 or 3.9% of cases had no detectable arvs or were missing data. Of those that self reported not on art, 24/306 or 7.8% had detectable arvs. Of those with missing artselfreported data, 64/410 or 16% had detectable arvs, and 1 respondent was missing arvstatus data. No respondents who self-reported aware and were tri90-unaware were included in this data because a respondent could not be self-reported aware and tri90-unaware based on the processing of the tri90aware variable. There were also no respondents with missing awareselfreported data that were classified as tri90-unaware. 
xtabs(~tri90aware + awareselfreported + artselfreported + arvstatus, bio_cong_dat) %>% 
  ftable(., col.vars = c("artselfreported", "arvstatus"))

# 60 respondents who had detectable arvs in their blood were reclassified as tri90aware from self-reported unaware or missing data. If someone did not have detectable arvs in their blood and were self-reported unaware, then they were tri90art-3 and not tri90aware. There were 87/714 or 12% of cases that were classified as on art even if they had no detectable arvs in their blood. Of those missing arvstatus data, 3/4 or 75% were classified as tri90art. These were likely self reported on art and missing arvstatus data. A relatively small number of respondents were missing arvstatus data. Of those with no detectable arvs, 369/714 were classified as tri90aware. Only if a respondent came back negative for arvstatus could they be considered tri90-unaware and only if a respondent was self-reported aware would the arvstatus be ignored when recoding the tri90aware variable. There were no respondents missing awareselfreported data in this dataset. There were no respondents who were missing arvstatus data that self-reported not on art.
xtabs(~tri90aware + awareselfreported + arvstatus + tri90art, bio_cong_dat) %>% 
  ftable(., col.vars = c("arvstatus", "tri90art"))

#This table shows that somehow 122 individuals with viral loads under 1000 copies/milliliter were classified as having viral loads over 200 copies/milliliter
xtabs(~vlunder200 + vls, bio_dat)
```

## "vlunder200"

The variable, vlunder200, was coded as ifelse(vlunder200 < 200, 1, 2), however, the variable was still a character when it ran. So, hundreds of observations had been misclassified. It should be fine now.

```{r viral load suppression data, eval = FALSE, include = FALSE}
bio_dat %>% 
  mutate(resultvlc = as.numeric(resultvlc, na.rm = TRUE)) %>% 
  filter(resultvlc < 201) # 2,165

# This doesn't work because the resultvlc hasn't been recoded
biomarker %>% 
  filter(hiv1statusfinalsurvey == 1) %>% 
  mutate(resultvlc = as.numeric(resultvlc, na.rm = TRUE)) %>% 
  filter(resultvlc < 201) # 392

# This shows 122 more people were viral load suppressed at 1000 copies per milliliter compared to 200 copies per milliliter. This is within the realm of what would be expected.
bio_dat %>% 
  filter(vls == 1) # 2,287

bio_cong_dat %>% 
  filter(vls == 1) # 2,242

bio_cong_dat %>% 
  filter(resultvlc < 200) # 2,011

#There are still 129 more folks who are virally suppressed at the 1000 copies per milliliter level, as expected
bio_cong_dat %>% 
  filter(vlunder200 == 1) # 2,011

#This does not make sense to me. This variable is not acting as expected.
bio_cong_dat %>% 
  filter(vlunder200 == 1,
         vls == 2) # 225
```

# Second 90 (arv status)

## If a person self reports that they are unaware and has detectable arvs in their blood, what is the probability of them being tri90aware?

## If a person self reports that they are unaware and is viral load suppressed (under 200 copies/ milliliter), what is the probability of them being tri90aware?

## If a person self reports that they are aware and has detectable arvs in their blood, what is the probability of them self-reporting that they are on arvs?

### We cannot test if a person has arvs in their blood nor if they self-report their arv status because they are not asked these questions if the respondent is self-reported unaware

# lca with pre-processed variables

```{r data for lca}
bio_dat_lca = bio_dat %>% 
  select(aware,
         art,
         awareselfreported, 
         arvstatus, 
         artselfreported, 
         vls, 
         tri90, 
         tri90aware, 
         tri90art, 
         tri90vls,
         vlunder200) %>% 
  mutate(aware = recode(aware, "1" = "1", "2" = "0", "99" = NULL),
         art = recode(art, "1" = "1", "2" = "0", "99" = NULL),
         awareselfreported = recode(awareselfreported, "1" = "1", "2" = "0", "99" = NULL),
         arvstatus = recode(arvstatus, "1" = "1", "2" = "0", "99" = NULL),
         artselfreported = recode(artselfreported, "1" = "1", "2" = "0", "99" = NULL),
         vls = recode(vls, "1" = "1", "2" = "0", "99" = NULL),
         tri90 = recode(tri90, "1" = "1", "2" = "0", "99" = NULL),
         tri90aware = recode(tri90aware, "1" = "1", "2" = "0", "99" = NULL),
         tri90art = recode(tri90art, "1" = "1", "2" = "0", "99" = NULL),
         tri90vls = recode(tri90vls, "1" = "1", "2" = "0", "99" = NULL),
         vlunder200 = recode(vlunder200, "1" = "1", "2" = "0", "99" = NULL)) 
```

```{r lca}
# full model
bio_lca_full = bio_dat_lca %>% 
  select(awareselfreported, arvstatus, vlunder200)

# simple model
# bio_lca_prep = bio_dat_lca %>%
#   select(awareselfreported)

set.seed(22)
bio_lca = randomLCA::randomLCA(bio_lca_full, calcSE = TRUE)
summary(bio_lca)

outcomeProbs(bio_lca)
BIC(bio_lca)
```

```{r nclasses full model, cache = TRUE}
outcome = NULL
for (i in 1:10) {
  outcome[[i]] = bio_dat_lca %>% 
    randomLCA(., freq=.$freq, nclass = i)
}

BIC = NULL
for(i in 1:10) {
  BIC[[i]] = BIC(outcome[[i]])
}

bic_data = NULL
for (i in 1:10) {
  bic_data[[i]] =  
    data.frame(nclasses = i,
               bic = BIC[[i]])
}

tibble(bic_data) %>% 
  unnest() %>% 
  knitr::kable()
```

# There are NAs in all the observations that are tri90aware == 0.

```{r model choosing, eval = FALSE}
# assume that tri90aware is the truth

bio_mod_lca = bio_dat_lca %>%
  mutate(across(.funs = as.numeric)) %>% 
  na.exclude()

nrow(bio_mod_lca)

bio_mod_lca %>% 
  group_by(tri90aware) %>% 
  summarize(sum = n())

bio_unr = bio_dat_lca %>% 
  filter(tri90aware == "0")

bio_r = bio_dat_lca %>% 
  filter(tri90aware == "1")

bio_mod = glm(tri90aware ~ . -aware -awareselfreported -tri90 -tri90aware, data = bio_mod_lca, family = poisson)

skimr::skim(bio_cong_dat)
skimr::skim(bio_mod_lca)
```


