---
title: "phia_lca4_jsg2145"
author: "Jared Garfinkel"
date: "8/21/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(randomLCA)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	cache = TRUE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## read in data

```{r}
biomarker <- read_csv(file = "./data/Shims22016adultbio.csv",
                      col_names = TRUE, 
                      col_types = NULL,
                      locale = default_locale(), 
                      na = c("", "NA"), 
                      quoted_na = TRUE,
                      quote = "\"", 
                      comment = "", 
                      trim_ws = TRUE, 
                      skip = 0,
                      n_max = Inf, 
                      progress = show_progress(), 
                      skip_empty_rows = TRUE)
```

```{r}
bio_dat <- biomarker %>% 
  filter(hiv1statusfinalsurvey == 1) %>%
  select(personid, 
         resultvlc) %>%
  mutate(resultvlc = recode(resultvlc, 
                    "< LLOD" = "1",
                    "< LLOQ: 20" = "20",
                    "< LLOQ: 40" = "40",
                    "< LLOQ: 400" = "400",
                    "< LLOQ: 839" = "839",
                    "> ULOQ 10000000" = "10000000"),
         resultvlc = as.numeric(resultvlc),
         vlunder200 = if_else(resultvlc < 200, 1, 2),
         across(.funs = as_factor))
```

The biomarker data set is `r nrow(bio_dat)` observations of `r ncol(bio_dat)` variables, "personid", "resultvlc", and "vlunder200".

It was filtered on hiv1statusfinalsurvey = 1.

```{r}
# Read public-release datasets
# Independent dataset
# hivstatusfinal not filtered
independent <- read_csv(file = "./data/Shims22016adultind.csv") %>% 
  filter(age > 14,
         bt_status == 1,
         hivstatusfinal == 1) 
```

After including only those who are 15 years old and over who had a blood test and were HIV+, the independent data set is `r nrow(independent)` observations of `r ncol(independent)` variables. We select only those related to the coding of "awareselfreported" variable and "personid."

```{r}
ind_dat_lca = independent %>%
  select(personid, known_hiv_status, hivtstever, hivtstrslt, hivpsbp, hivrtpg, hivrslr) %>% 
  tibble() %>% 
  mutate(across(.fns = as.numeric),
         across(everything(), ~recode(.x, "1" = 1, 
                                      "2" = 0, 
                                      "99" = 0, 
                                      "-8" = 0, 
                                      "3" = 0, 
                                      "4" = 0,
                                      "-9" = 0)),
         across(everything(), ~replace_na(.x, 0)))

# ind_dat_lca
# independent %>% 
#   filter(gender == 1) %>% 
#   nrow()
```

## suspect zero inflated data in pregnancy related questions (hivpsbp, hivrtpg, hivrslr)

Because the variables hivpsbp, hivrtpg and hivrslr are related to pregnancy, there are `r independent %>% filter(gender == 1) %>% nrow()` men respondents who are not asked these questions. This is problematic because these patients are not 0s, but ineligible for the outcome. This results in zero-inflated poisson data.

```{r}
skimr::skim(ind_dat_lca)
```

# lca

```{r}
set.seed(719)
ind_lca2 = randomLCA::randomLCA(ind_dat_lca, calcSE = TRUE)
lca_probs_ind2 = outcomeProbs(ind_lca2)
```

```{r}
set.seed(719)
aware_lca_r = randomLCA::randomLCA(ind_dat_lca, random = TRUE, calcSE = TRUE)
lca_probs_r = outcomeProbs(aware_lca_r, boot = TRUE)
# summary(aware_lca_r)
lca_probs_r
```

### extract classes

```{r}
set.seed(719)
lca_out_ind = postClassProbs(aware_lca_r) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 1, 0))
```

```{r}
ind_dat_out = ind_dat_lca %>% 
  left_join(lca_out_ind, copy = TRUE, by = c("known_hiv_status", "hivtstever", "hivtstrslt", "hivpsbp", "hivrtpg", "hivrslr"))
```

```{r}
xtabs(~hivpsbp + class, data = ind_dat_out)
```

There are `r ind_dat_out %>% filter(hivpsbp == 0, class == 1) %>% nrow()` false negatives in the hivpsbp variable. The sensitivity is therefore 346/2596, `r round(346/2596, 2)`.

# dealing with zero-inflated poisson data

## filter for women only??

This data only include women to reduce the zero inflation given that more women will have been asked about pregnancy. However, only a fraction of women have been pregnant.

```{r}
ind_w <- read_csv(file = "./data/Shims22016adultind.csv") %>% 
  filter(age > 14,
         bt_status == 1,
         hivstatusfinal == 1,
         gender == 2) %>% 
  select(known_hiv_status, hivtstever, hivtstrslt, hivpsbp, hivrtpg, hivrslr)
```

There are `r ind_w %>% nrow()` observations in our women only data set.

```{r}
ind_dat_w_lca = ind_w %>%
  tibble() %>% 
  mutate(across(.fns = as.numeric),
         across(everything(), ~recode(.x, "1" = 1, 
                                      "2" = 0, 
                                      "99" = 0, 
                                      "-8" = 0, 
                                      "3" = 0, 
                                      "4" = 0,
                                      "-9" = 0)),
         across(everything(), ~replace_na(.x, 0)))

skimr::skim(ind_dat_w_lca)
```

### run LCA without random effects

```{r}
set.seed(719)
ind_lca2w = randomLCA::randomLCA(ind_dat_w_lca, calcSE = TRUE)
lca_probs_ind2w = outcomeProbs(ind_lca2w)
```

### add random effects

```{r}
set.seed(719)
aware_lca_rw = randomLCA::randomLCA(ind_dat_w_lca, random = TRUE, calcSE = TRUE)
lca_probs_rw = outcomeProbs(aware_lca_rw, boot = TRUE)
# summary(aware_lca_r)
lca_probs_rw
```


### extract classes

```{r}
set.seed(719)
lca_out_ind_w = postClassProbs(aware_lca_rw) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 1, 0))
```

```{r}
ind_dat_out_w = ind_dat_w_lca %>% 
  left_join(lca_out_ind_w, copy = TRUE, by = c("known_hiv_status", "hivtstever", "hivtstrslt", "hivpsbp", "hivrtpg", "hivrslr"))
```

```{r}
xtabs(~hivpsbp + class, data = ind_dat_out_w)
```

#### There were `r ind_dat_out_w %>% filter(hivpsbp == 0, class == 1) %>% nrow()` false negatives in hivpsbp. The sensitivity of this test is 346/1816, `r round(346/1816, 2)`, slightly higher, but still practically uninformative.

### The same analysis without random effects

```{r}
set.seed(719)
lca_out_ind_w2 = postClassProbs(ind_lca2w) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 1, 0))
```

```{r}
ind_dat_out_w2 = ind_dat_w_lca %>% 
  left_join(lca_out_ind_w2, copy = TRUE, by = c("known_hiv_status", "hivtstever", "hivtstrslt", "hivpsbp", "hivrtpg", "hivrslr"))
```

```{r}
xtabs(~hivpsbp + class, data = ind_dat_out_w2)
```


## transform the pregnancy variables?

### There are three questions related to pregnancy included in this model. 

```{r}
pgyes = ind_dat_w_lca %>%
  filter(hivpsbp == 1 | hivrtpg == 1 | hivrslr == 1) # 437 rows
```

#### There are `r pgyes %>% nrow()` patients who had hiv results from one of these three variables. 

### We combine the three variables into one.

```{r}
ind_dat_w_lca2 = ind_dat_w_lca %>% 
  mutate(hivpg = if_else(hivpsbp == 1 | hivrtpg == 1 | hivrslr == 1, 1, 0)) %>% 
  select(-hivpsbp, -hivrtpg, -hivrslr)
```

```{r}
set.seed(719)
ind_lca3 = randomLCA::randomLCA(ind_dat_w_lca2, calcSE = TRUE)
lca_probs_ind3 = outcomeProbs(ind_lca3)
```

```{r}
set.seed(719)
aware_lca_r3 = randomLCA::randomLCA(ind_dat_w_lca2, random = TRUE, calcSE = TRUE)
lca_probs_r3 = outcomeProbs(aware_lca_r3, boot = TRUE)
# summary(aware_lca_r)
lca_probs_r3
```

### extract classes

```{r}
set.seed(719)
lca_out_ind3 = postClassProbs(aware_lca_r3) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 0, 1))
```

```{r}
ind_dat_out3 = ind_dat_w_lca2 %>% 
  left_join(lca_out_ind3, copy = TRUE, by = c("known_hiv_status", "hivtstever", "hivtstrslt", "hivpg"))
```

```{r}
xtabs(~hivpg + class, data = ind_dat_out3)
```

#### There are `r ind_dat_out3 %>% filter(hivpg == 0, class == 1) %>% nrow()` false negatives. The sensitivity is therefore 437/1816, `r round(437/1816, 2)`. Still better, but not useful.

## remove the pregnancy variables, add back vlunder200?

```{r}
full_df = left_join(bio_dat, independent, by = "personid")

full_df_lca = full_df %>% 
  mutate(across(.fns = as.numeric),
       across(everything(), ~recode(.x, "1" = 1, 
                                    "2" = 0, 
                                    "99" = 0, 
                                    "-8" = 0, 
                                    "3" = 0, 
                                    "4" = 0,
                                    "-9" = 0)),
       across(everything(), ~replace_na(.x, 0))) %>% 
  select(known_hiv_status, hivtstever, hivtstrslt, vlunder200)
```

#### Our full dataset has `r full_df_lca %>% nrow()` rows and `r full_df_lca %>% ncol()` variables. Three of these variables we keep from our previous model and 1 is added from the biomarker dataset. "vlunder200" is suspected to be a proxy for awareness because most of those who are aware will be on art in Eswatini.

### no random effects

```{r}
set.seed(719)
expanded_lca = randomLCA::randomLCA(full_df_lca, calcSE = TRUE)
lca_probs_ind2w = outcomeProbs(expanded_lca)
```

### include random effects

```{r}
set.seed(719)
aware_lca_full = randomLCA::randomLCA(full_df_lca, random = TRUE, calcSE = TRUE)
lca_probs_full = outcomeProbs(aware_lca_full, boot = TRUE)
# summary(aware_lca_r)
lca_probs_full
```


### extract classes

```{r}
set.seed(719)
lca_out_full = postClassProbs(aware_lca_full) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 0, 1))
```

```{r}
full_dat_out = full_df_lca %>% 
  left_join(lca_out_full, copy = TRUE, by = c("known_hiv_status", "hivtstever", "hivtstrslt", "vlunder200"))
```

```{r}
xtabs(~vlunder200 + class, data = full_dat_out)
```

#### There are `r full_dat_out %>% filter(vlunder200 == 0, class == 1) %>% nrow()` false negatives in "vlunder200". The sensitivity is 2059/2593, `r round(2059/2593, 2)`, a significant improvement. Including a 4th variable allows one to use the random effects option of the randomLCA package in R.

### extract classes without random effects

```{r}
set.seed(719)
lca_out_exp = postClassProbs(expanded_lca) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 0, 1))
```

```{r}
exp_dat_out = full_df_lca %>% 
  left_join(lca_out_exp, copy = TRUE, by = c("known_hiv_status", "hivtstever", "hivtstrslt", "vlunder200"))
```

```{r}
xtabs(~vlunder200 + class, data = exp_dat_out)
```
