---
title: "phia_lca_jsg2145"
author: "Jared Garfinkel"
date: "8/2/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(randomLCA)
library(pROC)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
# Read public-release datasets
#Biomarker dataset
biomarker <- read_csv(file = "./data/Shims22016adultbio.csv",
                      col_names = TRUE, 
                      col_types = NULL,
                      locale = default_locale(), 
                      na = c("", "NA"), 
                      quoted_na = TRUE,
                      quote = "\"", 
                      comment = "", 
                      trim_ws = TRUE, 
                      skip = 0,
                      n_max = Inf, 
                      progress = show_progress(), 
                      skip_empty_rows = TRUE)

#Individual interview dataset
individual <- read_csv(file = "./data/Shims22016adultind.csv",
                       col_names = TRUE, 
                       col_types = NULL,
                       locale = default_locale(), 
                       na = c("", "NA"), 
                       quoted_na = TRUE,
                       quote = "\"", 
                       comment = "", 
                       trim_ws = TRUE, 
                       skip = 0,
                       n_max = Inf, 
                       progress = show_progress(), 
                       skip_empty_rows = TRUE)
```

```{r clean data}
ind_dat <-
  filter(individual, bt_status == 1) %>% #
  select(personid, 
         gender, 
         age, 
         hivstatusfinal, 
         known_hiv_status, 
         arvscurrent) %>% 
  mutate(gender = factor(gender, 
                         labels = c("1" = "male",
                                    "2" = "female")),
         across(.funs = as_factor))

bio_dat <- biomarker %>% 
  filter(hiv1statusfinalsurvey == 1) %>%
  select(personid, 
         hiv1statusfinalsurvey,
         aware,
         art,
         awareselfreported, 
         arvstatus, 
         artselfreported, 
         resultvlc, 
         vls, 
         tri90, 
         tri90aware, 
         tri90art, 
         tri90vls) %>%
  mutate(resultvlc = recode(resultvlc, 
                    "< LLOD" = "1",
                    "< LLOQ: 20" = "20",
                    "< LLOQ: 40" = "40",
                    "< LLOQ: 400" = "400",
                    "< LLOQ: 839" = "839",
                    "> ULOQ 10000000" = "10000000"),
         resultvlc = as.numeric(resultvlc),
         vlunder200 = if_else(resultvlc < 200, 1, 2),
         across(.funs = as_factor))
```

```{r data for lca}
bio_dat_lca = bio_dat %>% 
  select(aware,
         art,
         awareselfreported, 
         arvstatus, 
         artselfreported, 
         vls, 
         tri90, 
         tri90aware, 
         tri90art, 
         tri90vls,
         vlunder200) %>% 
  mutate(aware = recode(aware, "1" = "0", "2" = "1", "99" = NULL),
         art = recode(art, "1" = "0", "2" = "1", "99" = NULL),
         awareselfreported = recode(awareselfreported, "1" = "0", "2" = "1", "99" = NULL),
         arvstatus = recode(arvstatus, "1" = "0", "2" = "1", "99" = NULL),
         artselfreported = recode(artselfreported, "1" = "0", "2" = "1", "99" = NULL),
         vls = recode(vls, "1" = "0", "2" = "1", "99" = NULL),
         tri90 = recode(tri90, "1" = "0", "2" = "1", "99" = NULL),
         tri90aware = recode(tri90aware, "1" = "0", "2" = "1", "99" = NULL),
         tri90art = recode(tri90art, "1" = "0", "2" = "1", "99" = NULL),
         tri90vls = recode(tri90vls, "1" = "0", "2" = "1", "99" = NULL),
         vlunder200 = recode(vlunder200, "1" = "0", "2" = "1", "99" = NULL))
```

```{r codebook}
# aware:                1 - Aware or considered aware because ARVs detectable 
#                       2 - Unaware and ARVs not detectable, or unaware and ARV testing results missing 
#                       99 - Missing
#
# art:                  1 - ARVs detectable, self-reported on ART, or both ARVs detectable and self-reported on ART 
#                       2 - Unaware or aware, ARVs not detectable and self-reported not on ART, or aware, missing ARV testing data and self-reported not on ART 
#                       99 - Missing
# awareselfreported:    1 - Self-report aware of HIV + status 
#                       2 - Self-report not aware of HIV + status 
#                       99 - missing
# artselfreported:      1 - On ART 
#                       2 - Not on ART 
#                       99 - Missing
# awareartselfreported: 1 - Self-report as not previously diagnosed 
#                       2 - Self-report as previously diagnosed, not on ART 
#                       3 - Previously diagnosed, on ART 
#                       99 - Missing, including incomplete tri90 information
# arvstatus:            1 - ARV detected 
#                       2 - ARV not detected 
#                       99 - Missing
# resultvlc:            > ULOQ 10000000 - Upper limit of quantification 10000000
#                       < LLOD - less than lower limit of detection 
#                       < LLOQ: 839 - less than lower limit of quantification of 839 
#                       < LLOQ: 400 - less than lower limit of quantification of 400
#                       < LLOQ: 40 - less than lower limit of quantification of 40 
#                       < LLOQ: 20 - less than lower limit of quantification of 20
# hivselfreport:        1 - Self-reported positive 
#                       2 - Self-reported negative 
#                       3 - Self-reported never tested or never received test result 
#                       99 - Missing
```

## First 90 (awareness)

```{r aware tables}


# This table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and had an art identified by a blood test
tbl1 = xtabs(~ aware + art, data = bio_dat)
chisq.test(tbl1)

# This contingency table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and the status of a dummy variable that measures whether they have suppressed viral loads
tbl2 = xtabs(~ aware + arvstatus, data = bio_dat)
chisq.test(tbl2)

# This contingency table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and reported that they were taking ART.
tbl3 = xtabs(~ aware + artselfreported, data = bio_dat)
chisq.test(tbl3)

# This contingency table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and the pre-processed self-reported awareness of their HIV status
tbl4 = xtabs(~ aware + awareselfreported, data = bio_dat)
chisq.test(tbl4)
```

```{r tri90aware tables}
# This table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and had an art identified by a blood test
tbl1 = xtabs(~ tri90aware + tri90art, data = bio_dat)
chisq.test(tbl1)

# This contingency table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and the status of a dummy variable that measures whether they have suppressed viral loads
tbl2 = xtabs(~ tri90aware + arvstatus, data = bio_dat)
chisq.test(tbl2)

# This contingency table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and reported that they were taking ART.
tbl3 = xtabs(~ tri90aware + artselfreported, data = bio_dat)
chisq.test(tbl3)

# This contingency table shows the respondents/informants who reported that they were aware of their hiv status or had an identifiable arv drug in their blood test and the pre-processed self-reported awareness of their HIV status
tbl4 = xtabs(~ tri90aware + awareselfreported, data = bio_dat)
chisq.test(tbl4)
```


```{r contingency table data}
bio_cong_dat = biomarker %>% 
  filter(tri90 == 1) %>% 
  select(personid, 
         hiv1statusfinalsurvey,
         aware,
         art,
         awareselfreported, 
         arvstatus, 
         artselfreported, 
         resultvlc, 
         vls, 
         tri90, 
         tri90aware, 
         tri90art, 
         tri90vls) %>%
  mutate(resultvlc = recode(resultvlc, 
                    "< LLOD" = "1",
                    "< LLOQ: 20" = "20",
                    "< LLOQ: 40" = "40",
                    "< LLOQ: 400" = "400",
                    "< LLOQ: 839" = "839",
                    "> ULOQ 10000000" = "10000000"),
         resultvlc = as.numeric(resultvlc),
         vlunder200 = if_else(resultvlc < 200, 1, 2),
         across(.funs = as_factor))
```

```{r}
nrow(bio_cong_dat)

# while all the respondents classified as unaware under tri90aware were unaware under awareselfreported, 59 respondents who self-reported they were unaware were reclassified as aware, and 1 respondent classified as missing data was reclassified as aware under tri90aware. A total of 60 respondents were reclassified from their self-reported aware status.
xtabs(~tri90aware + awareselfreported, bio_cong_dat)

# While no respondents classified as unaware under tri90aware were asked for artselfreported data, 306 of those who reported they were not on art were reclassified as aware and 64 respondents with missing artselfreported data were reclassified as aware
xtabs(~tri90aware + artselfreported, bio_cong_dat)

# While no respondents who tested positive for art were classified as aware, 282 respondents who tested negative for art were reclassified as aware, and 346 respondents who tested negative for art were classified as unaware.
xtabs(~ tri90aware + tri90art, data = bio_cong_dat)

# There were 736 respondents with viral load above 200 who were classified as aware. This may be indicative of treatment non-compliance. There were 96 respondents with viral load under 200 classified as unaware. This may indicate the presence of elite suppressors or transient viral load suppression among respondents.
xtabs(~tri90aware + vlunder200, bio_cong_dat)

# There were 369 respondents with no detectable arvs reclassified as tri90aware and 3 respondents with missing arvstatus data classified as aware. Meanwhile, 1 respondent with missing arvstatus data was classified as unaware. These 4 respondents missing arvstatus data may or may not have been reclassified.
xtabs(~tri90aware + arvstatus, bio_cong_dat)

# This table shows that 444 respondents with viral load above 1000 copies/ milliliter were reclassified as aware, while only 35 individuals with viral load under 1000 copies per milliliter were classified as unaware.
xtabs(~tri90aware + vls, bio_cong_dat)

# These variables are in agreement
xtabs(~tri90aware + aware, data = bio_cong_dat)

```

# Differential misclassification

Differential misclassification occurs when misclassification of exposure is not equal between subjects that have or do not have the health outcome, or when misclassification of the health outcome is not equal between exposed and unexposed subjects.

Assume the tri90 classifications represent the truth. It is possible to compare the classifications of respondents using pre-processed measurements (i.e. of blood samples and survey responses) to the "true" values and assess whether the misclassification is differential or non-differential.

```{r}
# The variable aware is useful here because it mirrors tri90aware, yet it shows that reclassification took place amont those self reported that they were unaware or were missing awareselfreported data. In fact, 59 respondents were reclassified from self-reportedly unaware to aware. None were reclassified from aware to unaware.
xtabs(~tri90aware + awareselfreported + aware, bio_cong_dat) %>% 
  ftable(.)

# This table shows that 4 respondents missing artselfreported data were classified as tri90aware who reported that they were aware. These individuals were asked whether they were on art, but the data is missing.
xtabs(~tri90aware + awareselfreported + artselfreported, bio_cong_dat) %>% 
  ftable(.)

# Some respondents marked that they were not on art and had no detectable arvs in their blood. All of them were classified as aware. All respondents who responded that they were not aware of their status were recoded as not on art because they were not aware of their status. 
xtabs(~tri90aware + awareselfreported + tri90art, bio_cong_dat) %>% 
  ftable(.)

# Of those with viral loads above 200 copies/milliliter, 720 reported that they were aware and were subsequently coded as tri90aware; 16 reported that they were unaware, but were ultimately coded as aware; 250 responded that they were unaware and were coded as unaware. So, 16/986 or 1.6% of those with elevated viral loads were reclassified. On the other hand, of those with suppressed viral loads, 43 reported they were unaware and were subsequently reclassified; 96 reported that they were unaware and found to be unaware. So, of those who self-reported that they were unaware, 96/405 or 24% had viral load under 200 copies/milliliter, and 59/405 or 15% were reclassified as tri90aware. 
xtabs(~tri90aware + awareselfreported + vlunder200, bio_cong_dat) %>% 
  ftable(.)

# There was 1 respondent with missing arvstatus information who selfreported that they were unaware and classified as unaware. None of those with missing arvstatus data were reclassified from their self-reported awareness. The same can be said of those who had no detectable arvs in their blood. Only those with detectable arvs were reclassified from unaware to aware.
xtabs(~tri90aware + awareselfreported + arvstatus, bio_cong_dat) %>% 
  ftable(.)

# This table shows that there were reclassifications based on vls in each category. 48 respondents were reclassified from self reported unaware to tri90aware who were viral load suppressed (less than 1000 copies/milliliter), while 11 who were not viral load suppressed (less than 1000 copies/milliliter) were also reclassified as tri90aware from self-reported unaware. One respondent who was not viral load suppressed was missing self-reported awareness data, but classified as tri90aware. Of those who were viral load suppressed 48/2242 or 2% were reclassified. Interestingly of those who self-reported that they were unaware, 83/405 or 20% were viral load suppressed (less than 1000 copies/milliliter). This number is smaller than vlunder200 data because missing data was recoded using the detection thresholds of the blood tests meaning that several respondents 
xtabs(~tri90aware + awareselfreported + vls, bio_cong_dat) %>% 
  ftable(.)

#This table shows that somehow 230 individuals with viral loads over 1000 copies/milliliter were classified as having viral loads under 200 copies/milliliter
xtabs(~vlunder200 + vls, bio_dat)
```

## "vlunder200"

The variable, vlunder200, was coded as ifelse(vlunder200 < 200, 1, 2), however, the variable was still a character when it ran. So, hundreds of observations had been misclassified. It should be fine now.

```{r viral load suppression data}
bio_dat %>% 
  mutate(resultvlc = as.numeric(resultvlc, na.rm = TRUE)) %>% 
  filter(resultvlc < 201) # 2,165

# This doesn't work because the resultvlc hasn't been recoded
biomarker %>% 
  filter(hiv1statusfinalsurvey == 1) %>% 
  mutate(resultvlc = as.numeric(resultvlc, na.rm = TRUE)) %>% 
  filter(resultvlc < 201) # 392

# This shows 122 more people were viral load suppressed at 1000 copies per milliliter compared to 200 copies per milliliter. This is within the realm of what would be expected.
bio_dat %>% 
  filter(vls == 1) # 2,287

bio_cong_dat %>% 
  filter(vls == 1) # 2,242

bio_cong_dat %>% 
  filter(resultvlc < 200) # 2,011

#There are still 129 more folks who are virally suppressed at the 1000 copies per milliliter level, as expected
bio_cong_dat %>% 
  filter(vlunder200 == 1) # 2,011

#This does not make sense to me. This variable is not acting as expected.
bio_cong_dat %>% 
  filter(vlunder200 == 1,
         vls == 2) # 225
```

