---
title: "phia_lca7_jsg2145"
author: "Jared Garfinkel"
date: "9/9/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(randomLCA)
library(gt)
library(caret)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	cache = TRUE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## read in data

```{r, echo = FALSE, results = 'hide', cache = FALSE, warning = FALSE}
# Load in PHIA data

recode_one_zero <- function(x) {
  if_else( is.na(x), 99,
  if_else( as.numeric(x) %in% c(0, 2), 0, # JSG
  if_else( as.numeric(x) == 1, 1, 99)
  ))
}

## read in data
bio <- read_csv("C:\\Users\\jared\\OneDrive\\Mailman\\Summer 2020\\PHIA ICAP\\zimphia2015adultbio.csv") %>% 
  janitor::clean_names() %>% 
  mutate(across(.cols = c(-country, -householdid, -personid, -centroidid, -surveystdt, -resultvlc), ~as.numeric(.x)))
ind <- read_csv("C:\\Users\\jared\\OneDrive\\Mailman\\Summer 2020\\PHIA ICAP\\zimphia2015adultind.csv") %>% 
  janitor::clean_names() %>% 
  mutate(across(.cols = c(-country, -householdid, -personid, -centroidid, -surveystdt), ~as.numeric(.x)))

bio_select <- bio %>% 
    left_join(ind, by = c("personid", "country", "householdid", "centroidid", "bt_status", "gender", "age", "surveystdt", "hivstatusfinal", "varstrat", "varunit")) %>% 
  # "avdprgmeth", "hivtstnors", "hivpostell", "sptgrgt", "stdl", "seekhelp", "unwantsxhelp")) %>%
    select(-contains("vmpstw"), -contains("hivk"), -contains("casi"), -contains("intw")) %>%
    filter(bt_status == 1,
           hivstatusfinal == 1) %>% 
    select(personid, age, gender, contains(c("status", "aware", "art", "vl", "tri90", "arv", "test", "hiv"))) %>% 
    mutate( resultvlc = recode(resultvlc, 
                               "ND" = "1",
                               "NQ: < 20" = "20",
                               "NQ: < 182" = "182"),
            resultvlc = as.numeric(resultvlc),
            vlunder200 = if_else(resultvlc < 200, 1, 2, missing = 2)) # JSG (from Stephen)

#table(vls = bio_select$vls, vl_lt_200 = bio_select$vl_lt_200)
# head(ind)
# head(bio)
# head(bio_select)
# str(ind$bt_status)
# bio$varstrat
# bio_select %>%
#   select(ends_with("y"))
# bio_select %>%
#   group_by(resultvlc) %>%
#   nest() %>%
#   filter(str_detect(resultvlc, "[:alpha:]"))
# which(is.na(bio_select$resultvlc == TRUE))
# bio_select$resultvlc
```

## make the variables, save awareartselfreported for later

```{r}
bio_select_mod <- bio_select %>%
    mutate( reported_positive = if_else( hivrslr == 1 | hivtstrslt == 1 | hivpsbp == 1 | hivrtpg == 1 | hivtfposy > 0, 1, 0, missing = 0),
            never_tested = if_else(hivtstever == 2, 1, 0, missing = 0),
            reported_negative = if_else(hivpsbp == 2 | (hivtsbp = 1 & hivrtpg == 2) | hivrslr == 2 | hivtstrslt == 2, 1, 0, missing = 0),
            pos_test_year = if_else(hivtfposy > 0, 1, 0),
            tested12mo = if_else(testedreceiveddetail %in% c(1,4), 1, 0),
            cd4testever = recode_one_zero(cd4testever),
            arvstakenev = if_else(arvstakenev == 1, 1, 0, missing = 0),
            arvscurrent = if_else(arvscurrent == 1, 1, 0, missing = 0),
            oth_art_care = if_else(arvtklb == 1 | arvcntn == 1 | hivcare == 1, 1, 0, missing = 0),
            across(-personid, recode_one_zero))

bio_select_var = bio %>% 
  filter(bt_status == 1,
         hivstatusfinal == 1) %>% 
  select(personid, awareartselfreported)

# nrow(bio_select_var) # 3507
```

## LCA for vlunder200-model, arvstatus-model, and simple-model

### vlunder200

```{r}
modelvars <- c( "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "vlunder200")
# Remove all rows which have values other than 0 or 1
aware_patterns <- bio_select_mod %>% 
  select( all_of(modelvars) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
awareARTVL.lca3 <- randomLCA(aware_patterns, nclass = 3, quadpoints = 30)

outcomep <- round(awareARTVL.lca3$outcomep, 4)
colnames(outcomep) <- names(aware_patterns)

OutcomeProbs.Table.AwareARTVL <- as.data.frame( cbind( Class = c("Class 1 - Aware, not on ART", "Class 2 - Aware, on ART", "Class 3 - Unaware"), outcomep)) %>%
  mutate(across(-Class, as.numeric))
```

### arvstatus

```{r}
modelvars_arv <- c( "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "arvstatus")
# Remove all rows which have values other than 0 or 1
aware_patterns_arv <- bio_select_mod %>% 
  select( all_of(modelvars_arv) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
awareART.lca3 <- randomLCA(aware_patterns_arv, nclass = 3, quadpoints = 30)

outcomep <- round(awareART.lca3$outcomep, 4)
colnames(outcomep) <- names(aware_patterns_arv)

OutcomeProbs.Table.AwareART <- as.data.frame( cbind( Class = c(1, 2, 3, 4), outcomep)) %>%
  mutate(across(-Class, as.numeric))
```

## extract classes

```{r}
set.seed(719)
lca_out_full = postClassProbs(awareARTVL.lca3) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3, 1,
                                 if_else(class_3 > class_1, 3, 1)), 
                         if_else(class_2 > class_3, 2, 3)),
         vlsaware = if_else(class == 1, 1, 0), # coded "forwards" meaning 1 is aware on art, 0 is EITHER unaware OR "aware and not on art"
         awarenotart = if_else(class == 2, 0, 1), # coded "backwards" meaning 0 is "aware not on art", 1 is not "aware not on art", or in other words, "1" is NEITHER unaware NOR aware and not on art - in terms of  class, a "1" means that an individual may be EITHER "aware AND on art" OR unaware
         unaware = if_else(class == 3, 0, 1)) # coded "backwards" meaning 0 is unaware, 1 is not unaware
```

```{r}
set.seed(719)
lca_out_full_arv = postClassProbs(awareART.lca3) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3, 1,
                                 if_else(class_3 > class_1, 3, 1)), 
                         if_else(class_2 > class_3, 2, 3)),
         vlsaware = if_else(class == 1, 1, 0), # coded "forwards" meaning 1 is aware on art, 0 is EITHER unaware OR "aware and not on art"
         awarenotart = if_else(class == 2, 0, 1), # coded "backwards" meaning 0 is "aware not on art", 1 is not "aware not on art", or in other words, "1" is NEITHER unaware NOR aware and not on art - in terms of  class, a "1" means that an individual may be EITHER "aware AND on art" OR unaware
         unaware = if_else(class == 3, 0, 1)) # coded "backwards" meaning 0 is unaware, 1 is not unaware
```


```{r}
full_dat_out = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_full, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "vlunder200"))

```

```{r}
full_dat_out_arv = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_full_arv, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "arvstatus"))
```

```{r}
class_out = full_join(full_dat_out, full_dat_out_arv, by = c("personid", "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "arvstatus", "vlunder200", "awareartselfreported.y"))
# nrow(class_out) # 3,507
# nrow(full_dat_out) #3,507
# nrow(full_dat_out_arv) # 3,507
```

# further tests, simple model, simpler model

## no lab data

```{r}
modelvars_simple <- c( "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care")
# Remove all rows which have values other than 0 or 1
aware_patterns_simple <- bio_select_mod %>% 
  select( all_of(modelvars_simple) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
aware.lca_simple <- randomLCA(aware_patterns_simple, nclass = 3, quadpoints = 30)

outcomep <- round(aware.lca_simple$outcomep, 4)
colnames(outcomep) <- names(aware_patterns_simple)

OutcomeProbs.Table.Aware <- as.data.frame( cbind( Class = c("Class 1 - Aware, not on ART", "Class 2 - Aware, on ART", "Class 3 - Unaware"), outcomep)) %>%
  mutate(across(-Class, as.numeric))
```

## extract classes

```{r}
set.seed(719)
lca_out_simple = postClassProbs(aware.lca_simple) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3, 1,
                                 if_else(class_3 > class_1, 3, 1)), 
                         if_else(class_2 > class_3, 2, 3)),
         vlsaware = if_else(class == 1, 1, 0), # coded "forwards" meaning 1 is aware on art, 0 is EITHER unaware OR "aware and not on art"
         awarenotart = if_else(class == 2, 0, 1), # coded "backwards" meaning 0 is "aware not on art", 1 is not "aware not on art", or in other words, "1" is NEITHER unaware NOR aware and not on art - in terms of  class, a "1" means that an individual may be EITHER "aware AND on art" OR unaware
         unaware = if_else(class == 3, 0, 1)) # coded "backwards" meaning 0 is unaware, 1 is not unaware
```

## append classes to dataframe

```{r}
simple_dat_out = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_simple, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care"))
```

## join simple dataframes with full dataframes

```{r}
simple_vls_class_out = full_join(full_dat_out, simple_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "arvstatus", "awareartselfreported.y"))

simple_arv_class_out = full_join(full_dat_out_arv, simple_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "awareartselfreported.y"))
```

## check the classes

```{r}
gt(OutcomeProbs.Table.AwareARTVL)
gt(OutcomeProbs.Table.AwareART)
gt(OutcomeProbs.Table.Aware)

```


```{r}
xtabs(~class.x + class.y, data = simple_vls_class_out)

xtabs(~class.x + class.y, data = simple_arv_class_out)
```

## awareartselfreported

```{r}
xtabs(~awareartselfreported.y + class, data = simple_dat_out, exclude = "99")

xtabs(~awareartselfreported.y + class, data = full_dat_out, exclude = "99")

xtabs(~awareartselfreported.y + class, data = full_dat_out_arv, exclude = "99")
```

## compare models to tri90 variables

### tri90aware vs. LCA model using arvstatus

```{r}
conmat_aware_df_arv = full_dat_out_arv

# 1 - aware on art
# 2 - aware not on art
# 3 - unaware

conmat_aware_df_arv$class = factor(conmat_aware_df_arv$class, levels = c("aware - on art" = 1, "aware - not on art" = 2, "unaware" = 3), labels = c("aware", "aware", "unaware"))
conmat_aware_df_arv$tri90aware = factor(conmat_aware_df_arv$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(conmat_aware_df_arv$class, conmat_aware_df_arv$tri90aware, positive = "aware")
cMat
```

```{r}
conmat_aware_df = full_dat_out

# unique(full_dat_out$arvstatus)

conmat_aware_df$class = factor(conmat_aware_df$class, levels = c("aware - not on art" = 1, "aware - on art" = 2, "unaware" = 3), labels = c("aware", "aware", "unaware"))
conmat_aware_df$tri90aware = factor(conmat_aware_df$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(conmat_aware_df$class, conmat_aware_df$tri90aware, positive = "aware")
cMat
```

```{r}
simple_conmat_df = simple_dat_out

simple_conmat_df$class = factor(simple_conmat_df$class, levels = c("aware - not on art" = 1, "aware - on art" = 2, "unaware" = 3), labels = c("aware", "aware", "unaware"))
simple_conmat_df$tri90aware = factor(simple_conmat_df$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(simple_conmat_df$class, simple_conmat_df$tri90aware, positive = "aware")
cMat

```

## tri90art

```{r}
conmat_aware_df_arv = full_dat_out_arv

# unique(full_dat_out_arv$arvstatus)

conmat_aware_df_arv$class = factor(conmat_aware_df_arv$class, levels = c("aware - on art" = 1, "aware - not on art" = 2, "unaware" = 3), labels = c("on art", "not on art", "not on art"))
conmat_aware_df_arv$tri90art = factor(conmat_aware_df_arv$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(conmat_aware_df_arv$class, conmat_aware_df_arv$tri90art, positive = "on art")
cMat
```

```{r}
conmat_aware_df = full_dat_out

# unique(full_dat_out$arvstatus)

conmat_aware_df$class = factor(conmat_aware_df$class, levels = c("aware - on art" = 1, "aware - not on art" = 2, "unaware" = 3), labels = c("on art", "not on art", "not on art"))
conmat_aware_df$tri90art = factor(conmat_aware_df$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(conmat_aware_df$class, conmat_aware_df$tri90art, positive = "on art")
cMat
```

```{r}
simple_conmat_df = simple_dat_out

simple_conmat_df$class = factor(simple_conmat_df$class, levels = c("aware - not on art" = 1, "aware - on art" = 2, "unaware" = 3), labels = c("not on art", "on art", "not on art"))
simple_conmat_df$tri90art = factor(simple_conmat_df$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(simple_conmat_df$class, simple_conmat_df$tri90art, positive = "on art")
cMat

```

# simpler model

## LCA, simpler

### vlunder200

```{r}
modelvars_simpler_vl <- c( "never_tested", "reported_negative", "reported_positive", "arvscurrent", "vlunder200")
# Remove all rows which have values other than 0 or 1
aware_patterns_simpler_vl <- bio_select_mod %>% 
  select( all_of(modelvars_simpler_vl) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
aware.lca_simpler_vl <- randomLCA(aware_patterns_simpler_vl, nclass = 3, quadpoints = 30)

outcomep <- round(aware.lca_simpler_vl$outcomep, 4)
colnames(outcomep) <- names(aware_patterns_simpler_vl)

OutcomeProbs.Table.Aware_simpler_vl <- as.data.frame( cbind( Class = c("Class 1 - Aware, not on ART", "Class 2 - Aware, on ART", "Class 3 - Unaware"), outcomep)) %>%
  mutate(across(-Class, as.numeric))
```

### arvstatus

```{r}
modelvars_simpler_arv <- c( "never_tested", "reported_negative", "reported_positive", "arvscurrent", "arvstatus")
# Remove all rows which have values other than 0 or 1
aware_patterns_simpler_arv <- bio_select_mod %>% 
  select( all_of(modelvars_simpler_arv) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
aware.lca_simpler_arv <- randomLCA(aware_patterns_simpler_arv, nclass = 4, quadpoints = 30)

outcomep_simpler_arv <- round(aware.lca_simpler_arv$outcomep, 4)
colnames(outcomep_simpler_arv) <- names(aware_patterns_simpler_arv)

OutcomeProbs.Table.Aware_simpler_arv <- as.data.frame( cbind( Class = c(1, 2, 3, 4), outcomep_simpler_arv)) %>%
  mutate(across(-Class, as.numeric))
```


## extract classes

```{r}
set.seed(719)
lca_out_simpler_vl = postClassProbs(aware.lca_simpler_vl) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3, 1,
                                 if_else(class_3 > class_1, 3, 1)), 
                         if_else(class_2 > class_3, 2, 3)))
         # vlsaware = if_else(class == 1, 1, 0), # coded "forwards" meaning 1 is aware on art, 0 is EITHER unaware OR "aware and not on art"
         # awarenotart = if_else(class == 2, 0, 1), # coded "backwards" meaning 0 is "aware not on art", 1 is not "aware not on art", or in other words, "1" is NEITHER unaware NOR aware and not on art - in terms of  class, a "1" means that an individual may be EITHER "aware AND on art" OR unaware
         # unaware = if_else(class == 3, 0, 1)) # coded "backwards" meaning 0 is unaware, 1 is not unaware
```

### What are the classes in the simpler vl model?

They don't seem to make much sense. Perhaps the first one is never tested, reported positive, not on arvs. The second one is reported negative, not on arvs, and the third one is reported positive, on arvs.

Maybe 2 classes would work better...

```{r}
gt(OutcomeProbs.Table.Aware_simpler_vl)
```

## simpler model with 4 classes

It appears 4 classes seems to do the trick...class 1 - never tested, class 2 - reported negative, class 3, aware not on art, class 4, aware on art.

### vlunder200

```{r}
modelvars_simpler_vl <- c( "never_tested", "reported_negative", "reported_positive", "arvscurrent", "vlunder200")
# Remove all rows which have values other than 0 or 1
aware_patterns_simpler_vl <- bio_select_mod %>% 
  select( all_of(modelvars_simpler_vl) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
aware.lca_simpler_vl2 <- randomLCA(aware_patterns_simpler_vl, nclass = 4, quadpoints = 30)

outcomep_svl2 <- round(aware.lca_simpler_vl2$outcomep, 4)
colnames(outcomep_svl2) <- names(aware_patterns_simpler_vl)

OutcomeProbs.Table.Aware_simpler_vl2 <- as.data.frame( cbind( Class = c(1, 2, 3, 4), outcomep_svl2)) %>%
  mutate(across(-Class, as.numeric))
gt(OutcomeProbs.Table.Aware_simpler_vl2)
```

## worth a try to add another class?

classes would apparently be: 1 - aware, not on art; 2 - reported negative; 3 - never tested; 4 - aware on art.

```{r}
modelvars <- c( "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "vlunder200")
# Remove all rows which have values other than 0 or 1
aware_patterns <- bio_select_mod %>% 
  select( all_of(modelvars) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
awareARTVL.lca4 <- randomLCA(aware_patterns, nclass = 4, quadpoints = 30)

outcomep4 <- round(awareARTVL.lca4$outcomep, 4)
colnames(outcomep4) <- names(aware_patterns)

OutcomeProbs.Table.AwareARTVL4 <- as.data.frame( cbind( Class = c(1, 2, 3, 4), outcomep4)) %>%
  mutate(across(-Class, as.numeric))
gt(OutcomeProbs.Table.AwareARTVL4)
```

For the arvstatus-model, classes could be viewed as: 1 - never tested; 2 - aware on art; 3 - aware not on art; 4 - reported negative

This might actually work...?

```{r}
modelvars <- c( "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "arvstatus")
# Remove all rows which have values other than 0 or 1
aware_patterns_arv4 <- bio_select_mod %>% 
  select( all_of(modelvars) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
awareART.lca4 <- randomLCA(aware_patterns_arv4, nclass = 4, quadpoints = 30)

outcomep_arv4 <- round(awareART.lca4$outcomep, 4)
colnames(outcomep_arv4) <- names(aware_patterns_arv4)

OutcomeProbs.Table.AwareART4 <- as.data.frame( cbind( Class = c(1, 2, 3, 4), outcomep_arv4)) %>%
  mutate(across(-Class, as.numeric))
gt(OutcomeProbs.Table.AwareART4)
```

For the simpler model (no lab work), the classes are: 1- aware on art; 2 - reported negative; 3 - aware not on art; and 4 - never tested.

Okay, so let's compare them.

```{r}
modelvars_simple <- c( "never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care")
# Remove all rows which have values other than 0 or 1
aware_patterns_simple4 <- bio_select_mod %>% 
  select( all_of(modelvars_simple) ) %>%
  filter( across(everything(), ~ .x %in% c(0,1)))
# aware_patterns$vlunder200
set.seed(719)
aware.lca_simple4 <- randomLCA(aware_patterns_simple, nclass = 4, quadpoints = 30)

outcomep_simple4 <- round(aware.lca_simple4$outcomep, 4)
colnames(outcomep_simple4) <- names(aware_patterns_simple4)

OutcomeProbs.Table.Aware4 <- as.data.frame( cbind( Class = c(1, 2, 3, 4), outcomep_simple4)) %>%
  mutate(across(-Class, as.numeric))
gt(OutcomeProbs.Table.Aware4)
```

## extract classes from the models with 4 LCA classes

```{r}
# simpler_vl2 - 4 variables + lab data
set.seed(719)
lca_out_simpler_vl2 = postClassProbs(aware.lca_simpler_vl2) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3,
                                 if_else(class_3 > class_4, 1,
                                         if_else(class_4 > class_1, 4, 1)),
                                 if_else(class_3 > class_1,
                                         if_else(class_1 > class_4, 3,
                                                 if_else(class_4 > class_3, 4, 3)),
                                         if_else(class_1 > class_4, 1, 4))),
                         if_else(class_2 > class_3, 
                                 if_else(class_3 > class_4, 2,
                                         if_else(class_4 > class_2, 4, 2)),
                                 if_else(class_4 > class_3, 4, 3))))

# simple - 5 SR variables, no lab data
set.seed(719)
lca_out_simple4 = postClassProbs(aware.lca_simple4) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3,
                                 if_else(class_3 > class_4, 1,
                                         if_else(class_4 > class_1, 4, 1)),
                                 if_else(class_3 > class_1,
                                         if_else(class_1 > class_4, 3,
                                                 if_else(class_4 > class_3, 4, 3)),
                                         if_else(class_1 > class_4, 1, 4))),
                         if_else(class_2 > class_3, 
                                 if_else(class_3 > class_4, 2,
                                         if_else(class_4 > class_2, 4, 2)),
                                 if_else(class_4 > class_3, 4, 3))))

# vlunder200 - full model (7 variables), vlunder200
set.seed(719)
lca_out_vls4 = postClassProbs(awareARTVL.lca4) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3,
                                 if_else(class_3 > class_4, 1,
                                         if_else(class_4 > class_1, 4, 1)),
                                 if_else(class_3 > class_1,
                                         if_else(class_1 > class_4, 3,
                                                 if_else(class_4 > class_3, 4, 3)),
                                         if_else(class_1 > class_4, 1, 4))),
                         if_else(class_2 > class_3, 
                                 if_else(class_3 > class_4, 2,
                                         if_else(class_4 > class_2, 4, 2)),
                                 if_else(class_4 > class_3, 4, 3))))

# arvstatus - full model (7 variables), arvstatus
set.seed(719)
lca_out_art4 = postClassProbs(awareART.lca4) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3,
                                 if_else(class_3 > class_4, 1,
                                         if_else(class_4 > class_1, 4, 1)),
                                 if_else(class_3 > class_1,
                                         if_else(class_1 > class_4, 3,
                                                 if_else(class_4 > class_3, 4, 3)),
                                         if_else(class_1 > class_4, 1, 4))),
                         if_else(class_2 > class_3, 
                                 if_else(class_3 > class_4, 2,
                                         if_else(class_4 > class_2, 4, 2)),
                                 if_else(class_4 > class_3, 4, 3))))

# simpler model (5 variables), arvstatus
set.seed(719)
lca_out_simpler_arv = postClassProbs(aware.lca_simpler_arv) %>% 
  janitor::clean_names() %>% 
  mutate(class = if_else(class_1 > class_2, 
                         if_else(class_2 > class_3,
                                 if_else(class_3 > class_4, 1,
                                         if_else(class_4 > class_1, 4, 1)),
                                 if_else(class_3 > class_1,
                                         if_else(class_1 > class_4, 3,
                                                 if_else(class_4 > class_3, 4, 3)),
                                         if_else(class_1 > class_4, 1, 4))),
                         if_else(class_2 > class_3, 
                                 if_else(class_3 > class_4, 2,
                                         if_else(class_4 > class_2, 4, 2)),
                                 if_else(class_4 > class_3, 4, 3))))
```

## compare to simple, vlunder200, arvstatus-models

```{r}
simpler_vl2_dat_out = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_simpler_vl2, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvscurrent", "vlunder200"))
```

```{r}
full_dat_out4 = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_vls4, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "vlunder200"))

```

```{r}
full_dat_out_arv4 = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_art4, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care", "arvstatus"))
```

```{r}
simple_out4 = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_simple4, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvstakenev", "arvscurrent", "oth_art_care"))
```

```{r}
simpler_vl2_vls_class_out = full_join(full_dat_out4, simpler_vl2_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "arvscurrent", "arvstatus", "awareartselfreported.y"))

simpler_vl2_arv_class_out = full_join(full_dat_out_arv4, simpler_vl2_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "vlunder200", "arvscurrent", "awareartselfreported.y"))

simpler_vl2_simple_class_out = full_join(simple_out4, simpler_vl2_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "vlunder200", "arvscurrent", "arvstatus", "awareartselfreported.y"))

vls_arv_class_out = full_join(full_dat_out4, full_dat_out_arv4, by = c("personid", "never_tested", "reported_negative", "reported_positive", "vlunder200", "arvstatus", "arvscurrent", "awareartselfreported.y"))
```

## cross tabulation

Let's check the full datasets first out of curiosity.

There is similar agreement between classes to the 3 class model.

```{r}
xtabs(~class.x + class.y, data = vls_arv_class_out)
```

What about other models? (comparing simplified models - no lab data, or fewer variables)

There are more disagreements between classes. But, how about against tri90 variables?

It seems much of the disagreement in classes is in class 3.

```{r}
xtabs(~class.x + class.y, data = simpler_vl2_vls_class_out)

xtabs(~class.x + class.y, data = simpler_vl2_arv_class_out)

xtabs(~class.x + class.y, data = simpler_vl2_simple_class_out)
```

## Reminder of class labels for each LCA output

```{r}
gt(OutcomeProbs.Table.Aware4)
gt(OutcomeProbs.Table.AwareART4)
gt(OutcomeProbs.Table.AwareARTVL4)
gt(OutcomeProbs.Table.Aware_simpler_vl2)
```

## fewer variables model, vlunder200

```{r}
simpler_vl2_conmat_df = simpler_vl2_dat_out

# 1 - never tested
# 2 - reported negative
# 3 - aware not on art
# 4 = aware on art

simpler_vl2_conmat_df$class = factor(simpler_vl2_conmat_df$class, levels = c("never tested" = 1, "reported negative" = 2, "aware not on art" = 3, "aware on art" = 4), labels = c("unaware", "unaware", "aware", "aware"))
simpler_vl2_conmat_df$tri90aware = factor(simpler_vl2_conmat_df$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(simpler_vl2_conmat_df$class, simpler_vl2_conmat_df$tri90aware, positive = "aware")
cMat
```

```{r}
simpler_vl2_conmat_df = simpler_vl2_dat_out

# 1 - never tested
# 2 - reported negative
# 3 - aware not on art
# 4 = aware on art

simpler_vl2_conmat_df$class = factor(simpler_vl2_conmat_df$class, levels = c("never tested" = 1, "reported negative" = 2, "aware not on art" = 3, "aware on art" = 4), labels = c("not on art", "not on art", "not on art", "on art"))
simpler_vl2_conmat_df$tri90art = factor(simpler_vl2_conmat_df$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(simpler_vl2_conmat_df$class, simpler_vl2_conmat_df$tri90art, positive = "on art")
cMat
```

## no lab data model

```{r}
simple_out4_conmat_df = simple_out4

# 1 - aware on art
# 2 - reported negative
# 3 - aware not on art
# 4 = never tested

simple_out4_conmat_df$class = factor(simple_out4_conmat_df$class, levels = c("aware on art" = 1, "reported negative" = 2, "aware not on art" = 3, "never tested" = 4), labels = c("on art", "not on art", "not on art", "not on art"))
simple_out4_conmat_df$tri90art = factor(simple_out4_conmat_df$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(simple_out4_conmat_df$class, simple_out4_conmat_df$tri90art, positive = "on art")
cMat
```

```{r}
simple_out4_conmat_df = simple_out4

# 1 - aware on art
# 2 - reported negative
# 3 - aware not on art
# 4 = never tested

simple_out4_conmat_df$class = factor(simple_out4_conmat_df$class, levels = c("aware on art" = 1, "reported negative" = 2, "aware not on art" = 3, "never tested" = 4), labels = c("aware", "unaware", "aware", "unaware"))
simple_out4_conmat_df$tri90aware = factor(simple_out4_conmat_df$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(simple_out4_conmat_df$class, simple_out4_conmat_df$tri90aware, positive = "aware")
cMat
```

## vlunder200 full model

```{r}
full_dat_out4_conmat_df = full_dat_out4

# 1 - aware not on art
# 2 - reported negative
# 3 - never tested
# 4 = aware on art

full_dat_out4_conmat_df$class = factor(full_dat_out4_conmat_df$class, levels = c("aware not on art" = 1, "reported negative" = 2, "never tested" = 3, "aware on art" = 4), labels = c("aware", "unaware", "unaware", "aware"))
full_dat_out4_conmat_df$tri90aware = factor(full_dat_out4_conmat_df$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(full_dat_out4_conmat_df$class, full_dat_out4_conmat_df$tri90aware, positive = "aware")
cMat
```

```{r}
full_dat_out4_conmat_df = full_dat_out4

# 1 - aware not on art
# 2 - reported negative
# 3 - never tested
# 4 = aware on art

full_dat_out4_conmat_df$class = factor(full_dat_out4_conmat_df$class, levels = c("aware not on art" = 1, "reported negative" = 2, "never tested" = 3, "aware on art" = 4), labels = c("not on art", "not on art", "not on art", "on art"))
full_dat_out4_conmat_df$tri90art = factor(full_dat_out4_conmat_df$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(full_dat_out4_conmat_df$class, full_dat_out4_conmat_df$tri90art, positive = "on art")
cMat
```

## arvstatus model

```{r}
full_dat_arv4_conmat_df = full_dat_out_arv4

# 1 - never tested
# 2 - aware on art
# 3 - aware not on art
# 4 = reported negative

full_dat_arv4_conmat_df$class = factor(full_dat_arv4_conmat_df$class, levels = c("never tested" = 1, "aware on art" = 2, "aware not on art" = 3, "reported negative" = 4), labels = c("unaware", "aware", "aware", "unaware"))
full_dat_arv4_conmat_df$tri90aware = factor(full_dat_arv4_conmat_df$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(full_dat_arv4_conmat_df$class, full_dat_arv4_conmat_df$tri90aware, positive = "aware")
cMat
```

```{r}
full_dat_arv4_conmat_df = full_dat_out_arv4

# 1 - never tested
# 2 - aware on art
# 3 - aware not on art
# 4 = reported negative

full_dat_arv4_conmat_df$class = factor(full_dat_arv4_conmat_df$class, levels = c("never tested" = 1, "aware on art" = 2, "aware not on art" = 3, "reported negative" = 4), labels = c("not on art", "on art", "not on art", "not on art"))
full_dat_arv4_conmat_df$tri90art = factor(full_dat_arv4_conmat_df$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(full_dat_arv4_conmat_df$class, full_dat_arv4_conmat_df$tri90art, positive = "on art")
cMat
```

## now with the simpler arvstatus model

```{r}
simpler_arv_dat_out = bio_select_mod %>% 
  left_join(bio_select_var, by = "personid") %>% 
  left_join(lca_out_simpler_arv, copy = TRUE, by = c("never_tested", "reported_negative", "reported_positive", "arvscurrent", "arvstatus"))
```


```{r}
simpler_arv_full_class_out = full_join(full_dat_out4, simpler_arv_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "arvscurrent", "arvstatus", "awareartselfreported.y"))

simpler_arv_arv_class_out = full_join(full_dat_out_arv4, simpler_arv_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "vlunder200", "arvscurrent", "awareartselfreported.y"))

simpler_arv_simple_class_out = full_join(simple_out4, simpler_arv_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "vlunder200", "arvscurrent", "arvstatus", "awareartselfreported.y"))

simpler_vls_arv_class_out = full_join(simpler_vl2_dat_out, simpler_arv_dat_out, by = c("personid", "never_tested", "reported_negative", "reported_positive", "vlunder200", "arvstatus", "arvscurrent", "awareartselfreported.y"))
```

## simple models compared to each other

Some disagreement. How about the other models?

```{r}
xtabs(~class.x + class.y, data = simpler_vls_arv_class_out)
```

Disagreement in class 4, sometimes class 3.

```{r}
xtabs(~class.x + class.y, data = simpler_arv_full_class_out)

xtabs(~class.x + class.y, data = simpler_arv_arv_class_out)

xtabs(~class.x + class.y, data = simpler_arv_simple_class_out)
```

## compare to tri90 variables

### tri90aware

#### cross-tabulation

reminder of classes

```{r}
gt(OutcomeProbs.Table.Aware_simpler_arv)
```


```{r}
simpler_arv_conmat_df = simpler_arv_dat_out

# 1 - aware on art
# 2 - never tested
# 3 - reported negative
# 4 - aware not on art

simpler_arv_conmat_df$class = factor(simpler_arv_conmat_df$class, levels = c("aware on art" = 1, "never tested" = 2, "reported negative" = 3, "aware not on art" = 4), labels = c("on art", "not on art", "not on art", "not on art"))
simpler_arv_conmat_df$tri90art = factor(simpler_arv_conmat_df$tri90art, c("on art" = 1, "not on art" = 0), labels = c("on art", "not on art"))
cMat = caret::confusionMatrix(simpler_arv_conmat_df$class, simpler_arv_conmat_df$tri90art, positive = "on art")
cMat
```

```{r}
simpler_arv_conmat_df = simpler_arv_dat_out

# 1 - aware on art
# 2 - never tested
# 3 - reported negative
# 4 - aware not on art

simpler_arv_conmat_df$class = factor(simpler_arv_conmat_df$class, levels = c("aware on art" = 1, "never tested" = 2, "reported negative" = 3, "aware not on art" = 4), labels = c("aware", "unaware", "unaware", "aware"))
simpler_arv_conmat_df$tri90aware = factor(simpler_arv_conmat_df$tri90aware, c("aware" = 1, "unaware" = 0), labels = c("aware", "unaware"))
cMat = caret::confusionMatrix(simpler_arv_conmat_df$class, simpler_arv_conmat_df$tri90aware, positive = "aware")
cMat
```

NOTE: For some reason the vlunder200 full model performs slightly better than the arvstatus full model in the tri90art variable.

However, in a simpler model, the arvstatus simpler model performs better than the vlunder200 simpler model.

The major improvement is from the simpler model to the full model in the tri90art accuracy.