---
title: "Xia et al. Bayesian LCA - Reproduction and Extension"
author: "Giles Reid"
date: "2/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# Load libraries:
library(tidyverse)
library(colorspace)
library(googledrive)
library(rjags)
library(runjags)
```

# Reproducing the Eswatini analysis

The Xia et al. (2020) paper used the `rjags` to fit a Bayesian latent class model to PHIA data from four countries. In the below we reproduce the Eswatini analysis and try out some extensions to this.

## 1. Load data for Eswatini

```{r}
# ?drive_auth_configure
# drive_auth_configure()
# drive_auth_config(active = FALSE)
# drive_auth(email = "jsg2145@icap.columbia.edu")

# drive_auth(email = "jsg2145@icap.columbia.edu", scopes = "https://www.googleapis.com/auth/drive.readonly")

drive_auth(scopes = "https://www.googleapis.com/auth/drive.readonly")

drive_user()

# x <- drive_get("~/PHIA Data Collaboration/Data/Public Release/Eswatini/Shims22016adultbio.csv")
# x <- drive_get("https://docs.google.com/spreadsheets/d/1FtvuLal09-fpY0awf759dCSoimueItKF/view?usp=sharing")
x <- as_id("https://drive.google.com/file/d/1FtvuLal09-fpY0awf759dCSoimueItKF/view?usp=sharing")
y <- as_id("https://drive.google.com/file/d/1FNfNzuQRcUWWDEnzcjYCzJjRgzKdrie2/view?usp=sharing")
# x$id
temp <- tempfile()
biodl <- drive_download(file = x, path = temp, overwrite = TRUE)
temp <- tempfile()
inddl <- drive_download(file = y, path = temp, overwrite = TRUE)
# str(bio)
bio <- read_csv(biodl$local_path) %>% 
  janitor::clean_names() %>% 
  mutate(across(.cols = c(varstrat, varunit), .fns = as.numeric))
ind <- read_csv(inddl$local_path)
head(bio)
head(ind)
```

## Select HIV+ people with non-missing values of the relevant variables

```{r}
eswatini.data <- ind %>%
  select(-contains("intwt"), -contains("vmpstw"), -contains("rfirs"), -starts_with("part")) %>%
  mutate(bt_status = as.numeric(bt_status)) %>% 
  left_join(bio, by = c("country", "householdid", "personid", "gender", "age", "hivstatusfinal", "surveystdt", "bt_status", "varstrat", "varunit")) %>% 
# 
# head(eswatini.data) %>% 
#   select(ends_with(".x"))
  filter(bt_status == 1 & hivstatusfinal == 1 & arvstatus %in% c(1,2) & awareselfreported %in% c(1,2))
```

## Definitions of the categories for analysis

Xia defines the following groups, for two different outcome variables, awareness of HIV+ status and 'ever tested' status, as follows:

| Group | True/unobserved state  | ARV test result | Self-reported status |
|-------|------------------------|-----------------|----------------------|
| A     | Truly aware/tested     | ARV+            | aware/tested         |
| B     | Truly unaware/untested | ARV+            | aware/tested         |
| C     | Truly aware/tested     | ARV-            | aware/tested         |
| D     | Truly unaware/untested | ARV-            | aware/tested         |
| E     | Truly aware/tested     | ARV+            | unaware/untested     |
| F     | Truly unaware/untested | ARV+            | unaware/untested     |
| G     | Truly aware/tested     | ARV-            | unaware/untested     |
| H     | Truly unaware/untested | ARV-            | unaware/untested     |

The population is split into observed groups as follows:

$$
y_1 = A + B \\
y_2 = C + D \\
y_3 = E + F \\
y_4 = G + H \\
$$The inputs to the model are just these four counts, $y_1 - y_4$. Coding these using our data:

```{r}
model.counts <- eswatini.data %>%
  mutate( y.category =
    case_when(
      arvstatus == 1 & awareselfreported == 1 ~ 1,
      arvstatus == 2 & awareselfreported == 1 ~ 2,
      arvstatus == 1 & awareselfreported == 2 ~ 3,
      arvstatus == 2 & awareselfreported == 2 ~ 4,
      TRUE ~ 99
    )
  ) %>%
  filter( y.category != 99 ) %>%
  select( personid, y.category, age, gender, awareselfreported, arvstatus, hivstatusfinal ) %>%
  group_by(awareselfreported, arvstatus, y.category) %>%
  summarise(count = n())

model.counts
```

To recap, these $y$s count the following groups of PLHIV:

-   $y_1$: ARV+ and self-reported aware,
-   $y_2$: ARV- and self-reported aware,
-   $y_3$: ARV+ and self-reported unaware,
-   $y_4$: ARV- and self-reported unaware.

## Setting up the model

The `rjags` package connects to the JAGS ("Just Another Gibbs Sampler") package, which runs outside of R and needs to be installed separately. To use it, we have to have a model definition. For the model in Xia et al.'s paper, I started from the equations in the supplemental materials, and translated these into the BUGS language. These equations define how to calculate the likelihood for the model, given a set of parameter values. They need to be written in such a way that they can be tracked back to a constant or a sample-able probability distribution, so that the Gibbs sampler/MCMC algorithm can proceed through its update/accept cycle. The full encoded model looks like this:

    model {

      p[1] = pi * sesrpos * covarv             + (1 - pi) * (1 - spsr) * (1 - sparv)
      p[2] = pi * sesrneg * (1 - covarv)       + (1 - pi) * (1 - spsr) * sparv
      p[3] = pi * (1 - sesrpos) * covarv       + (1 - pi) * spsr * (1 - sparv)
      p[4] = pi * (1 - sesrneg) * (1 - covarv) + (1 - pi) * spsr * sparv
      sesrneg = 1 - ratio * (1 - sesrpos)
      
      y[1:4] ~ dmulti(p[1:4], N)
      pi ~ dunif(0, 1)
      sesrpos ~ dunif(0, 1)
      spsr = 1
      
      covarv ~ dunif(0, 1)  
      sparv = 1
      R ~ dnorm( 0.3920421, 4.0)
      ratio = exp( R )
      
    }

This is saved in an external text file and becomes an input to the `jags.model` function. I've also put in some initial values, this is maybe not strictly necessary but helps give a more robust result.

### Explanation of the model code

The variables are as follows:

+--------------+--------------------------------------------------------------------------+-----------------------------------------------------+
| Name in code | Description                                                              | Value                                               |
+==============+==========================================================================+=====================================================+
| `sesrpos`    | Sensitivity of self-reporting among ARV+ people                          | Uniform random number in [0, 1]                     |
+--------------+--------------------------------------------------------------------------+-----------------------------------------------------+
| `sesrpos`    | Sensitivity of self-reporting among ARV- people                          | Uniform random number in [0, 1]                     |
+--------------+--------------------------------------------------------------------------+-----------------------------------------------------+
| `spsr`       | Specificity of self-reported status                                      | Assumed to be 1                                     |
+--------------+--------------------------------------------------------------------------+-----------------------------------------------------+
| `covarv`     | Proportion of PLHIV with detectable ARVs                                 | Uniform random number in [0, 1]                     |
+--------------+--------------------------------------------------------------------------+-----------------------------------------------------+
| `sparv`      | Specificity of lab ARV status                                            | Assumed to be 1                                     |
+--------------+--------------------------------------------------------------------------+-----------------------------------------------------+
| `ratio`      | Ratio of non-disclosure for lab ARV- PLHIV, compared with lab ARV+ PLHIV | Exponential of a normal distribution defined by R   |
+--------------+--------------------------------------------------------------------------+-----------------------------------------------------+
| `R`          | Normal distribution used to compute `ratio`                              | Normal distribution with mean `log(1.49)` and sd 4. |
+--------------+--------------------------------------------------------------------------+-----------------------------------------------------+
| `pi`         | The true proportion of individuals who are aware                         | Uniform random number in [0, 1]                     |
+--------------+--------------------------------------------------------------------------+-----------------------------------------------------+
| `y[1-4]`     | Counts of each category $y_1 - y_4$                                      | Calculated from the `p` values                      |
+--------------+--------------------------------------------------------------------------+-----------------------------------------------------+
| `p[1-4]`     | Membership probabilities in each category $y_1 - y_4$                    | Computed from the other parameters                  |
+--------------+--------------------------------------------------------------------------+-----------------------------------------------------+

So there's no restriction or prior on the sensitivities or the proportions with ARVs or truly aware. The `ratio` parameter is probably the most interesting one - it's set here based on other studies showing that people are less likely/more reluctant to say they know they're HIV+ when they aren't on treatment. The distribution here looks like this:

```{r ratio plot, message = FALSE, warning = FALSE}

ratio_vals <- as_tibble_col( exp( rnorm(10000, mean = log(1.48), 4.0)), "ratio" )
ggplot(ratio_vals, aes(x = ratio)) + 
  geom_histogram() + 
  scale_x_continuous(limits = c(0, 10)) + 
  scale_y_continuous(limits = c(0, 1000)) + theme_bw()

```

So it has a mean of around 1.5, and most values are between about 1 and 2.5. More extreme values are still allowed by this distribution, however.

## Running the model

Now that we have the data and the model, we can use `rjags` to actually fit the model, via the `jags.model`, `update`, and `coda.samples` functions. The inputs to `jags.model` are the counts computed from the PHIA data, the model file, and some initial values, plus some optional parameters for the model. The initial values aren't strictly required but help ensure a robust result.

`jags.model` creates the model object ready to be sampled from, `update` runs through a specified number of iterations for burn-in, and `coda.samples` does the final sampling from the model runs, saving values of the variables specified into an output file for analysis.

```{r rjags, cache = TRUE}
model.data <- list( "y" = model.counts$count, "N" = sum(model.counts$count))

initial.values <- list(
                        list( pi = 0.8, R = 0.0, covarv = .6),
                        list( pi = 0.6, R = 0.4, covarv = .8),
                        list( pi = 0.9, R = 0.3, covarv = .7),
                        list( pi = 0.3, R = 0.1, covarv = .2)
                      )

model <- jags.model("./data/xiamodel.txt", data = model.data, n.chains = 4)

update(model, n.iter=1000)
samples <- coda.samples(model, variable.names=c("pi", "ratio", "sesrpos", "sesrneg", "covarv", "p"),
                        n.iter=1000)

summary(samples)
plot(samples)

(fitted.stats <- bind_cols(var = rownames(summary(samples)$statistics), as_tibble(summary(samples)$statistics)))
(fitted.quantiles <- bind_cols(var = rownames(summary(samples)$statistics), as_tibble(summary(samples)$quantiles)))
```

These calculations all come out the same as what's reported in the Xia paper for Eswatini, so we seem to have set up the model correctly. You can see some of the standard model fitting plots here too - I've suppressed these for the rest of the document for brevity.

There's also another package called `run.jags` which can be used - it has some other options and outputs but the basic results should be the same.

```{r run.jags, cache = TRUE}
monitors <- c("pi", "ratio", "sesrpos", "sesrneg", "covarv")

out <- run.jags(model="./data/xiamodel.txt", monitor=monitors,
                 data=model.data, n.chains=4, method="parallel", inits=initial.values)

print(out)
summary(out)
#plot(out)

fitted.values <- bind_cols(var = monitors, as_tibble(summary(out)))
```

## What if we use viral load less than 200 instead of ARV status?

Because the model explicitly computes an estimate (`pi`) of the true awareness among the PLHIV population, we can use this same model to compute an adjusted prevalence using other variables. For example, here's what happens if we use viral load less than 200 copies/mL instead of the ARV status. Note here that the assumption of perfect specificity of ARVs, which seems very reasonable in that case (ie. nobody is going to have detected ARVs who isn't truly aware and on ARVs), is perhaps not quite as robust due to the possible existence of people who are 'elite controllers' (virally suppressed without ARVs).

```{r vltesting, cache = TRUE}
model.counts.vl <- eswatini.data %>%
  mutate(  resultvln = as.numeric(resultvlc),
           vl_lt_200 = case_when(
                        grepl("<", resultvlc, fixed = TRUE) ~ 1,
                        resultvln >= 0 & resultvln < 200 ~ 1,
                        resultvln >= 200 ~ 2,
                        TRUE ~ 99),
            y.category =
            case_when(
              vl_lt_200 == 1 & awareselfreported == 1 ~ 1,
              vl_lt_200 == 2 & awareselfreported == 1 ~ 2,
              vl_lt_200 == 1 & awareselfreported == 2 ~ 3,
              vl_lt_200 == 2 & awareselfreported == 2 ~ 4,
              TRUE ~ 99
            )
  ) %>%
  filter( y.category != 99 ) %>%
  select( personid, y.category, age, gender, awareselfreported, arvstatus, vl_lt_200, hivstatusfinal ) %>%
  group_by(awareselfreported, vl_lt_200, y.category) %>%
  summarise(count = n())

model.counts.vl

model.data.vl <- list( "y" = model.counts.vl$count, "N" = sum(model.counts.vl$count))

initial.values <- list(
  list( pi = 0.8, R = 0.0, covarv = .6),
  list( pi = 0.6, R = 0.4, covarv = .8)
)

model.vl <- jags.model("./data/xiamodel.txt", data = model.data.vl, n.chains = 2)

update(model.vl, n.iter=1000)
samples.vl <- coda.samples(model.vl, variable.names=c("pi", "ratio", "sesrpos", "sesrneg", "covarv", "p"),
                        n.iter=1000)

#plot(samples.vl)
```

Let's compare the results using the viral load variable to ARV status:

```{r comparison}

summary(samples.vl)
summary(samples)

```

Using viral load the estimate of the true value of awareness is 89.5% (88.1 - 91.2), compared to the estimate using ARVs of 89.08% (87.8 - 90.5), in other words extremely similar, but with a slightly higher standard error.

## Multi-country analysis

I ran similar analysis across all the PHIA countries (except Zambia because the dataset was not loading properly). Here's some graphs of the main results.

First a comparison of the unadjusted proportion aware (`UnAdjWgtP`, just using self-reported awareness), the adjusted awareness (`AdjWgtP`, using the standard PHIA recode with arvstatus), and the Bayesian LCA model estimated true awareness, using arvstatus to adjust (`arvstatus`).

```{r multicountry graphs 1}

library(pals)
library(RColorBrewer)
library(colorRamps)

z = as_id("https://drive.google.com/file/d/1bZ0-ShZ8535-3d7GvOYkiMLllbb5e58S/view?usp=sharing")

multicountry.datadl <- drive_download(file = z, path = temp, overwrite = TRUE)

# str(multicountry.datadl)

multicountry.data = readxl::read_xlsx(multicountry.datadl$local_path)
#awareness <- 
  #filter(var == "covarv")
  #filter(var == "ratio")
  #filter(var == "sesrpos")
  #filter(var == "sesrneg")

pd <- position_dodge(width = 0.5)

multicountry.data %>%
  filter(var == "pi" & !(model %in% c("AdjWgtP", "UnAdjWgtP", "vl_lt_200")) ) %>%
  ggplot(aes(x = country, y = Mean, colour = model, group = model)) +
  geom_point(position = pd, size = 2) +
  geom_errorbar(aes(ymin = Lower95, ymax = Upper95), width = 0.2, position = pd) +
  expand_limits(y = c(0.5, 1)) +
  theme_bw() +
  ylab("Estimated % aware")

```

We can also compare the adjusted awareness from using the arvstatus PHIA recode to the model using arvstatus and the model using viral load < 200 copies/mL (`vl_lt_200`).

```{r multicountry2}


multicountry.data %>%
  filter(var == "pi" & (model %in% c("AdjUnwgtP", "arvstatus", "vl_lt_200")) ) %>%
  ggplot(aes(x = country, y = Mean, colour = model, group = model)) +
  geom_point(position = pd, size = 2) +
  geom_errorbar(aes(ymin = Lower95, ymax = Upper95), width = 0.2, position = pd) +
  expand_limits(y = c(0.5, 1)) +
  theme_bw() +
  ylab("Estimated % aware")

```

In some countries the estimate is much lower using the viral load. What's going on here? We can examine the estimates for the proportion of self-reported aware and self-reported unawares who have viral load under 200, as well as the sensitivity estimates in each model:


```{r multicountry3}

multicountry.data %>%
  filter(var == "covarv" & (model %in% c("arvstatus", "vl_lt_200")) ) %>%
  ggplot(aes(x = country, y = Mean, colour = model, group = model)) +
  geom_point(position = pd, size = 2) +
  geom_errorbar(aes(ymin = Lower95, ymax = Upper95), width = 0.2, position = pd) +
  expand_limits(y = c(0.5, 1)) +
  theme_bw() +
  ylab("Estimated % of PLHIV with ARVs/low VL")


multicountry.data %>%
  filter(var == "sesrpos" & (model %in% c("arvstatus", "vl_lt_200")) ) %>%
  ggplot(aes(x = country, y = Mean, colour = model, group = model)) +
  geom_point(position = pd, size = 2) +
  geom_errorbar(aes(ymin = Lower95, ymax = Upper95), width = 0.2, position = pd) +
  expand_limits(y = c(0.5, 1)) +
  theme_bw() +
  ylab("Sensitivity of self-reported awareness among those WITHOUT ARVs/low vl")

multicountry.data %>%
  filter(var == "sesrneg" & (model %in% c("arvstatus", "vl_lt_200")) ) %>%
  ggplot(aes(x = country, y = Mean, colour = model, group = model)) +
  geom_point(position = pd, size = 2) +
  geom_errorbar(aes(ymin = Lower95, ymax = Upper95), width = 0.2, position = pd) +
  expand_limits(y = c(0.5, 1)) +
  theme_bw() +
  ylab("Sensitivity of self-reported awareness among those WITH ARVs/low vl")

```


Also we can look at the probabilities to be in each class:


```{r multicountry4}

multicountry.data %>%
  filter(grepl("p[", var, fixed = TRUE) & (model %in% c("arvstatus", "vl_lt_200")) ) %>%
  ggplot(aes(x = country, y = Mean, colour = var, shape = model)) +
  geom_point(position = pd, size = 2) +
  geom_errorbar(aes(ymin = Lower95, ymax = Upper95), width = 0.2, position = pd) +
  expand_limits(y = c(0.5, 1)) +
  theme_bw() +
  ylab("Probability of class membership")
```

In at least a few countries, $p_4$ is a lot higher than $p_1$ for the `vl_lt_200` model. Class one is the self-reported aware people with ARVs/low vl detected, and class four is the self-report unaware with ARVs/low vl not detected.


```{r}
multicountry.data %>%
  filter(var == "sesrpos" & (model %in% c("arvstatus", "vl_lt_200")) ) %>%
  ggplot(aes(x = country, y = Mean, colour = model, group = model)) +
  geom_point(position = pd, size = 2) +
  geom_errorbar(aes(ymin = Lower95, ymax = Upper95), width = 0.2, position = pd) +
  expand_limits(y = c(0.5, 1)) +
  theme_bw() +
  ylab("Sensitivity of self-reported awareness among those WITHOUT ARVs/low vl")

multicountry.data %>%
  filter(var == "sesrneg" & (model %in% c("arvstatus", "vl_lt_200")) ) %>%
  ggplot(aes(x = country, y = Mean, colour = model, group = model)) +
  geom_point(position = pd, size = 2) +
  geom_errorbar(aes(ymin = Lower95, ymax = Upper95), width = 0.2, position = pd) +
  expand_limits(y = c(0.5, 1)) +
  theme_bw() +
  ylab("Sensitivity of self-reported awareness among those WITH ARVs/low vl")

```



```{r multicountry111}
# Plot to show the magnitude of adjustments
aware_adj_plot <- multicountry.data %>% 
  select(country, var, Mean, model, Lower95, Upper95) %>%
  pivot_wider(id_cols = c(country, var), names_from = model, values_from = c(Mean, Lower95, Upper95)) %>%
  filter(var == "pi")


ggplot(data = aware_adj_plot, aes(x = Mean_AdjUnwgtP, y = Mean_arvstatus, colour = country, group = country)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = Lower95_arvstatus, ymax = Upper95_arvstatus)) +
  geom_abline(slope = 1, intercept = 0) +
  expand_limits(y = c(0.4, 1), x = c(0.4, 1)) +
  theme_bw() +
  scale_colour_manual(values = colorRampPalette(brewer.pal(8, "Dark2"))(12))


```